---
title: "Swedish eel silvering model summary"
subtitle: "SWEELHT"
author: "Viktor Thunell" 
date: "`r format(Sys.time(), '%d %B, %Y')`"
format: 
  html:
    code-fold: true
    code-summary: "Show code"
    page-layout: full
    embed-resources: true
    toc: true
knitr: 
  opts_chunk:
    fig.align: center
    out-width: 80%
editor: source
execute: 
  echo: true
  eval: true
  cache: true
---

# Latent classs logistic regression for eel migration probability

## Load libraries

```{r}
#| message: false
#| warning: false
#| cache: false

# Load libraries, install if needed
pkgs <- c("tidyverse", "tidylog", "devtools","viridis","sf","patchwork","posterior","ggridges", "tidybayes","bayesplot","nimble","here","reshape2","sdmTMB","broom")

if(length(setdiff(pkgs,rownames(installed.packages()))) > 0){
    install.packages(setdiff(pkgs, rownames(installed.packages())), dependencies = T)
  }

invisible(lapply(pkgs, library, character.only = T))

devtools::source_url("https://raw.githubusercontent.com/VThunell/diasp-lht/main/R/functions/map-plot.R")
options(ggplot2.continuous.colour = "viridis")
theme_set(theme_minimal()) 

# Set path
home <- here::here()
```

## Read data

```{r}
#| message: false
#| warning: false
#| cache: false

## Read data
#sweels <- readRDS(file = paste0(home, "/data/sweel_data_2026-06-24.RData"))
## Read data for 4f
sweels <- readRDS(file = paste0(home, "/data/sweel_data_2026-08-12.RData"))
```

```{r}
# lets keep the blue ones as females:
sweels %>%
  mutate(ld = if_else(length > 650 & is.na(sex),"pf","na")) %>%
  ggplot() +
  geom_point(aes(length, weight, color = ld)) +
  facet_wrap(~sex)
```

## Silvering logistic regression model

### Define data

```{r}
# filter females and keep only weights and lengths
df.sweel <- sweels %>%
  filter(!is.na(length)) %>%
  mutate(weight = if_else(weight == 0, NA, weight)) %>%
  # remove rows where used variables have at least two non NAs
  filter(rowSums(!is.na(select(., visual_stage, weight, length, age))) >= 2) %>%
  #keep the more than 50 cm long NA sex as females
  mutate(sex = if_else(is.na(sex) & length > 650, "female", sex)) %>%
  filter(sex == "female")
# make the con,eye and pec index 

abw <- df.sweel %>%
  drop_na(weight,length) %>%
  lm(data = ., log(weight) ~ log(length)) %>%
  tidy()

abe <- df.sweel %>%
  drop_na(diameter_eye,length) %>%
  lm(data = ., log(diameter_eye) ~ log(length)) %>%
  tidy()

abp <- df.sweel %>%
  drop_na(length_pectoral,length) %>%
  lm(data = ., log(length_pectoral) ~ log(length)) %>%
  tidy()

 
df.sweel2 <- df.sweel %>% 
  mutate(silver_vis = case_when(visual_stage == "Blankål" ~ 1,
                                visual_stage == "Halvblank ål" ~ NA,
                                visual_stage %in% c("Gulål","Yngel","Juvenil") ~ 0,
                                .default = NA),
         # silver_dur = case_when(durif_stage %in% c("FIV","FV") ~ 1,
         #                        durif_stage %in% c("FIII") ~ NA,
         #                        durif_stage %in% c("I","FII","MII") ~ 0,
         #                        .default = NA),
         length_sc = (length - mean(length, na.rm = TRUE)) / sd(length, na.rm = TRUE),
         age_sc = (age - mean(age, na.rm = TRUE)) / sd(age, na.rm = TRUE),
         con_index = exp(abw$estimate[1] + abw$estimate[2]*log(length))/weight,
         pec_index = exp(abp$estimate[1] + abp$estimate[2]*log(length))/length_pectoral,
         eye_index = exp(abe$estimate[1] + abe$estimate[2]*log(length))/diameter_eye,
         con_index_sc = (con_index - mean(con_index, na.rm = TRUE)) / sd(con_index, na.rm = TRUE),
         pec_index_sc = (pec_index - mean(pec_index, na.rm = TRUE)) / sd(pec_index, na.rm = TRUE),
         eye_index_sc = (eye_index - mean(eye_index, na.rm = TRUE)) / sd(eye_index, na.rm = TRUE),
         sb = case_when(sea_basin == "VH" ~ 1,
                        !sea_basin == "VH" ~ 2),
         # create dummy su for NA basins
         #spat.unit = if_else(is.na(spat.unit),str_c("dum_",sb),spat.unit),
         #spat.unit = if_else(marine == "yes", sb, main_catch),
         sa = as.integer(factor(sea_area)),
         spat.unit = if_else(marine == "yes", sa, main_catch),
         su = as.integer(factor(spat.unit)),
         #sb = as.integer(factor(sea_basin)),
         habitat = if_else(marine == "yes", 1,0),
         lat_sc = (lat - mean(lat, na.rm = TRUE)) / sd(lat, na.rm = TRUE),
         )  %>%
  # join in pectoral and eye index where rows only in res ( 8,717) are
  #left_join(res) %>%
  #mutate(silvratio = mean(silver_vis == 1, na.rm = TRUE), .by = spat.unit) %>%
  mutate(silver_vis_orig = silver_vis,
    #balance = if_else(silvratio < 0.1 | silvratio > 0.9 | is.na(silvratio), 0, 1),
         # the dummy su:s (Östersjön and Västerhavet) should only contribute to the sb level and thus be unbalanced
         #balance = if_else(str_detect(spat.unit, "dum_"), 0, balance),
         ind_id = row_number())

mask_to_na <- function(df, ratio = 8) {
  set.seed(1)
  df %>%
    group_by(su) %>%
    group_modify(~ {
      n_min <- sum(.x$silver_vis == 0, na.rm = TRUE)
      n_maj <- sum(.x$silver_vis == 1, na.rm = TRUE)
      
      if (n_maj > ratio * n_min & n_min > 0) {
        # mask excess majority class
        maj_rows <- which(.x$silver_vis == 1)
        mask_rows <- sample(maj_rows, size = n_maj - round(ratio * n_min))
        .x$silver_vis[mask_rows] <- NA
        .x } else if (n_min > ratio * n_maj & n_maj > 0) {
        # mask excess minority class
        min_rows <- which(.x$silver_vis == 0)
        mask_rows <- sample(min_rows, size = n_min - round(ratio * n_maj))
        .x$silver_vis[mask_rows] <- NA
        .x } else {
          .x
        }
      }) %>%
    ungroup()
}


df.sweel3 <- df.sweel2 %>%
#  mask_to_na() %>%
  mutate(silvratio = mean(silver_vis == 1, na.rm = TRUE), .by = spat.unit) %>%
  mutate(balance = if_else(silvratio < 0.1 | silvratio > 0.9 | is.na(silvratio), 0, 1)) 

# seems to work as intende, unbalanced units become balanced unless the ratio is 0/1
df.sweel3 %>% 
   mutate(silvratio_orig = mean(silver_vis_orig == 1, na.rm = TRUE), .by = spat.unit) %>%
  distinct(su,silvratio_orig,silvratio,balance)

df.sweel2 %>%
  summarise(n(), .by = silver_vis)

# the sb:s have data so we can get sb level nested hierarchy
df.sweel3 %>%
  filter(!(is.na(silver_vis))) %>%
  rowwise() %>%
  mutate(silver = max(c(silver_vis), na.rm = TRUE)) %>%
  ungroup() %>%
  summarise(n = n(), .by = c(silver,su)) %>%
  pivot_wider(id_cols = c(su), names_from = silver, values_from = n)

df.sweel3 %>%
  ggplot() +
  geom_histogram(aes(length, fill = factor(silver_vis)), color = NA, alpha = 0.3) +
  facet_wrap(~Lokal)
  
df.sweel3 %>%
  summarise(m = mean(length), n(), .by = c(sea_basin,silver_vis)) %>% arrange(sea_basin)
  summarise(n(), .by = c(sea_basin,silver_vis)) %>% arrange(sea_basin)
  # filter(spat.unit == 2120032080,
  #        silver_vis ==1) %>%
  # pull(length) %>%
  # summary()

df.sweel3 %>%
  select(length_sc, age_sc, con_index_sc, pec_index_sc, eye_index_sc) %>%
  drop_na(length_sc, age_sc, con_index_sc, pec_index_sc, eye_index_sc) %>%
  ggpairs(
  lower = list(continuous = wrap("smooth", alpha = 0.3, color = "blue")), # Scatter plots with trendline
  diag = list(continuous = wrap("densityDiag", fill = "gray")),           # Density plots on diagonal
  upper = list(continuous = wrap("cor", size = 4), progress = FALSE)
) 

```

### Plot silvering data
```{r}
#In space
plot_map_Nord +
  #geom_sf(data = ca_sf %>% filter(main_catch %in% df_cobaca_inland$main_catch), aes(fill = factor(main_catch)), color = NA,  alpha = 0.3)  +
  geom_point(data = df.sweel3 %>%
               mutate(Habitat = if_else(marine == "no", "Inland", "Marine")) %>%
               #filter(spat.unit == "Ö Sydkustens kustvatten" | is.na(spat.unit)) %>%
               add_utm_columns(ll_names = c("lon", "lat"), utm_crs = 32633), aes(X*1000, Y*1000, color = Habitat), size = 0.1) +
  guides(color = guide_legend(title.position = "top")) +
  labs(color = "") +
  theme(legend.position = "inside", legend.position.inside = c(0.18, 0.80))

ggsave(paste0(home, "/figures/map_data.pdf"), width = 18, height = 13, units = "cm")

# length at age by stage and sex
df.sweel3 %>%
  #ggplot(aes(age, length)) +
  ggplot(aes(age, length, color = factor(sb))) +
  geom_point(size = 0.1, alpha = 0.3) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5), show.legend = FALSE) 
  facet_wrap(~visual_stage)
  
# relationship between age, length and silver
df.sweel3 %>%
  ggplot(aes(length, silver_vis)) +
  geom_point(size = 0.1, alpha = 0.3) +
  facet_wrap(~su)
# df.sweel3 %>%
#   mutate(silv = silver_vis * silver_dur,
#          silv = if_else(is.na(silv), 0.5, silv)) %>%
#   ggplot(aes(silv, age)) +
#   geom_point(size = 0.1, alpha = 0.3) +
#   geom_smooth(method = "lm", se = TRUE) 
  
# relationship between length and le_cren, latitude
df.sweel3 %>%
  ggplot(aes(con_index_sc, length, color = factor(silver_vis))) +
  geom_point(size = 0.1, alpha = 0.3) +
  #geom_smooth(method = "gam", formula = y ~ s(x, k = 3), show.legend = FALSE) +
  geom_smooth(method = "lm", se = TRUE) +

df.sweel3 %>%
  ggplot(aes(eye_index_sc, length, color = factor(silver_vis))) +
  geom_point(size = 0.1, alpha = 0.3) +
  geom_smooth(method = "lm", se = TRUE) +
  facet_wrap(~habitat)
  
  
df.sweel3 %>%
  ggplot(aes(pec_index_sc, length, color = factor(silver_vis))) +
  geom_point(size = 0.1, alpha = 0.3) +
  #geom_smooth(method = "gam", formula = y ~ s(x, k = ), show.legend = FALSE) +
  geom_smooth(method = "lm", se = TRUE) +

df.sweel3 %>%
  ggplot(aes(lat_sc, length, color = factor(silver_vis))) +
  geom_point(size = 0.1, alpha = 0.3) +
  #geom_smooth(method = "gam", formula = y ~ s(x, k = 3), show.legend = FALSE) +
  geom_smooth(method = "lm", se = TRUE) 


# relationship between length and silver
df.sweel3 %>%
  mutate(silv = silver_vis * silver_dur,
         silv = if_else(is.na(silv), 0.5, silv)) %>%
  ggplot(aes(silv, length)) +
  geom_point(size = 0.1, alpha = 0.3) +
  geom_smooth(method = "lm", se = TRUE) +
  theme_minimal()

# relationship between length and silver
df.sweel3 %>%
  mutate(silv = silver_vis * silver_dur,
         silv = if_else(is.na(silv), 0.5, silv)) %>%
  ggplot() +
  geom_histogram(aes(length, fill = factor(tu)), alpha = 0.5) +
  facet_wrap(~silv)

# length of the silvers (where one of the classifiers says it is silver)
df.sweel3 %>%
  mutate(silver_dur = if_else(is.na(silver_dur),-1,silver_dur),
         silver_vis = if_else(is.na(silver_vis),-1,silver_vis),
         silv = silver_dur+silver_vis) %>%
  filter(!silv < 0) %>%
  ggplot(aes(length, factor(su), color = db)) +
  geom_boxplot(outlier.color = "grey",
               outlier.size = 0.5) 

# LHTs of Durif staged eels
df.sweel3 %>%
  filter(!is.na(durif_stage)) %>%
  ggplot(aes(length, factor(su), color = durif_stage)) +
  geom_boxplot(outlier.color = "grey",
               outlier.size = 0.5) 
df.sweel3 %>%
  filter(!is.na(durif_stage)) %>%
  ggplot(aes(age, factor(su), color = durif_stage)) +
  geom_boxplot(outlier.color = "grey",
               outlier.size = 0.5) 
df.sweel3 %>%
  filter(!is.na(durif_stage)) %>%
  ggplot(aes(weight1, factor(su), color = durif_stage)) +
  geom_boxplot(outlier.color = "grey",
               outlier.size = 0.5) 

# length of visual staged eels
df.sweel3 %>%
  filter(!is.na(visual_stage)) %>%
  ggplot(aes(length, factor(su), color = visual_stage)) +
  geom_boxplot(outlier.color = "grey",
               outlier.size = 0.5) 

# length-weight
df.sweel3 %>%
  drop_na(length,weight1) %>%
  ggplot(aes(length, weight1,color = factor(silver_dur))) +
  geom_point(size = 0.1, alpha = 0.3) 

# numbers of inds per clssifier    
df.sweel3 %>%
  summarise(n = n(), .by = c(visual_stage,durif_stage)) %>%
  arrange(visual_stage,-n)

df.sweel3 %>%
  filter(!(is.na(silver_vis) & is.na(silver_dur))) %>%
  rowwise() %>%
  mutate(silver = max(c(silver_vis,silver_dur), na.rm = TRUE)) %>%
  ungroup() %>%
  summarise(n = n(), .by = c(silver,spat.unit)) %>%
  pivot_wider(id_cols = spat.unit, names_from = silver, values_from = n) %>%
  arrange(-`1`,spat.unit)

df.sweel3 %>%
  filter(!(is.na(silver_vis) & is.na(silver_dur))) %>%
  rowwise() %>%
  mutate(silver = max(c(silver_vis,silver_dur), na.rm = TRUE)) %>%
  ungroup() %>%
  summarise(n = n(), .by = c(silver,sea_basin)) %>%
  pivot_wider(id_cols = sea_basin, names_from = silver, values_from = n) %>%
  arrange(-`1`,spat.unit)
```
### n over space and time 

```{r}
df.sweel3 %>%
  summarise(count = n(), .by = c(su))

df.sweel3 %>%
  summarise(count = n(), .by = c(year, su)) %>%
  ggplot(aes(year, factor(su), fill = count)) +
  geom_tile(color = "gray30") +
  #geom_vline(xintercept = 1998, color = "darkgrey") +
  scale_fill_viridis_c() +
  labs(y = "", x = "year") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Source models / Load samples
```{r, cache = FALSE}
# Load samples
sweelsilv.samples <- readRDS(paste0(home,"/data/samples/sweelsilv.samples_e2026-06-24.RData"))
sweelsilv.samples <- readRDS(paste0(home,"/data/samples/sweelsilv.samples_d2026-06-26.RData"))
sweelsilv.samples <- readRDS(paste0(home,"/data/samples/sweelsilv.samples_f2026-07-02.RData"))
sweelsilv.samples <- readRDS(paste0(home,"/data/samples/sweelsilv.samples_f2026-08-10.RData"))
sweelsilv.samples <- readRDS(paste0(home,"/data/samples/sweelsilv.samples_g2026-07-30.RData"))
sweelsilv.samples <- readRDS(paste0(home,"/data/samples/sweelsilv.samples_g2026-08-11.RData"))
sweelsilv.samples <- readRDS(paste0(home,"/data/samples/sweelsilv.samples_g_BALANCE_by_NAs2026-08-14.RData"))
sweelsilv.samples <- readRDS(paste0(home,"/data/samples/sweelsilv.samples_i2026-08-18.RData"))

sweelsilv.samples$WAIC
# for simple 33029.01 - 20260507
# for spatial alpha 13956.79 - 20260507
# for spatial and hierachical alpha 13875.92 - 20260508
# for spatial b0 and hierachical spatial alpha 13570.31 - 20260508
# for spatial and hierachical alpha and b0 13506.59 - 20260511
# for spatial and hierachical alpha and b0 21417.19 - 20260511 (this adds data so incomparable to previous models)
# for spatial and hierachical alpha and b0 with two classifiers and se/sp: 8411.084
# 225708 # new data and model after meeting 4/6 (model 4, no condition)
#  234143.4 20260615 and then 231746.8 for the 4c model with 
# 232228.8 4c
# 231643.5 4e
# 231671 4e 22june
# 44273.07 4d 22june
# f 1419142 3 july
# f 23/7: 165149.1
# g fixed and balanced 286095.4
# g fixed and balanced, all coastal sites 284486.4
# g more fixed and balanced,  4722275
s.samples <- sweelsilv.samples$samples
# # Or source model code and mcmc (load libs and data)
# t <- Sys.time()
# source(file = paste0(home,"/models_eel/model_eel_silvering.R"))
# Sys.time() - t # approx 4.5 hours with 10000 hmc samples
# WAIC based on combined log prob of silver_vis and _dur; d: 25569.66 (22946.08), e:26438.2
#401627.6 vs
ll_vis <- s.samples[, grep("logProb_silver_vis", colnames(s.samples))]
ll_dur <- s.samples[, grep("logProb_silver_dur", colnames(s.samples))]
ll_total <- ll_vis + ll_dur  # n_draws x n_individuals

# Compute WAIC manually
# lppd    <- sum(log(colMeans(exp(ll_total))))
# p_waic  <- sum(apply(ll_total, 2, var))
# #WAIC
# -2 * (lppd - p_waic)
# 
lppd    <- sum(log(colMeans(exp(ll_vis))))
p_waic  <- sum(apply(ll_vis, 2, var))
#WAIC
-2 * (lppd - p_waic)

```

### Traces, ess, and ‘potential scale reduction factor’

Variables "s.sig","Ustar","U","mu.a.gl","sd.a.gl","alpha","bs","bt")
```{r, cache = FALSE}
variable.names(as_draws(s.samples)) %>%
  str_remove("\\[.*\\]") %>% unique()

node.sub <- subset_draws(as_draws(s.samples), variable = 
                           c("mu_a_gl","mu_b_gl"))
                           #c("alpha_sb","logb0_sb","mu_a_gl","mu_b_gl"))
node.sub[] %>% mcmc_trace() 
node.sub %>% summarise_draws(c("mean","median","sd","rhat","ess_bulk","ess_tail"))
#log odds intercept to prob 1/(1+exp(z)). 
node.sub %>% summarise_draws(c("mean"))
1/(1+exp(-15)) # a silver of mean length in the data (631 mm) has a 99,99 % prob to be a silver

node.sub <- subset_draws(as_draws(s.samples), variable = c("alpha"))
#node.sub[,,sample(1:dim(node.sub)[3], 16)] %>% mcmc_trace() # two chain
node.sub[,sample(1:dim(node.sub)[2], 16)] %>% mcmc_trace() # one chain
node.sub %>% summarise_draws(c("mean","median","sd","rhat","ess_bulk","ess_tail")) %>% arrange(-rhat)

node.sub <- subset_draws(as_draws(s.samples), variable = c("logb0"))
node.sub[,sample(1:dim(node.sub)[2], 16)] %>% mcmc_trace() # one chain
node.sub %>% summarise_draws(c("mean","median","sd","rhat","ess_bulk","ess_tail")) %>% arrange(-rhat) %>% mutate(mean = exp(mean))

node.sub <- subset_draws(as_draws(s.samples), variable = #c("sp_v","sp_d","se_v","se_d","b1","b2","b3","b4","ab","w_nu","w_sd"))
c("sp_v","se_v","bA","bC","bLa","bH","bP","bE"))
node.sub[] %>% mcmc_trace() # one chain
node.sub %>% summarise_draws(c("mean","median","sd","rhat","ess_bulk","ess_tail"))
  
node.sub <- subset_draws(as_draws(s.samples), variable = "p")
#node.sub[,,sample(1:dim(node.sub)[3], 16)] %>% mcmc_trace()
node.sub[,sample(1:dim(node.sub)[2], 36)] %>% mcmc_trace()
#node.sub %>% mcmc_trace() # one chain
node.sub[,sample(1:dim(node.sub)[2], 600)] %>% summarise_draws(c("median","sd","rhat","ess_bulk","ess_tail")) %>% arrange(-rhat)

node.sub <- subset_draws(as_draws(s.samples), variable = "silver_rep")
node.sub <- subset_draws(as_draws(s.samples), variable = "measured_rep")
node.sub[,sample(1:dim(node.sub)[2], 36)] %>% mcmc_trace()
#node.sub %>% mcmc_trace() # one chain
node.sub[,sample(1:dim(node.sub)[2], 600)] %>% summarise_draws(c("median","sd","rhat","ess_bulk","ess_tail")) %>% arrange(-rhat)

```

### Silvering variables

```{r}
# alpha_sb/beta0_sb (sea basin unit), baltic vs west coast
# s.samples %>%
#   as_draws() %>%
#   spread_draws(alpha[su]) %>%
#   ggplot() +
#   geom_density(aes(x = alpha, color = factor(su))) +
# 
# s.samples %>%
#   as_draws() %>%
#   spread_draws(logb0[sb]) %>%
#   ggplot() +
#   geom_density(aes(x = logb0_sb, color = factor(sb))) 

# alpha an b0 by su
s.samples %>%
  as_draws() %>%
  spread_draws(alpha[su]) %>%
  left_join(df.sweel3 %>% distinct(su,sb,marine)) %>%
  ungroup() %>%
  ggplot() +
  geom_boxplot(aes(x = alpha, y = su, color = factor(sb), group = factor(su)), outliers = FALSE) +

s.samples %>%
  as_draws() %>%
  spread_draws(logb0[su]) %>%
  left_join(df.sweel3 %>% distinct(su,sb,marine)) %>%
  ggplot() +
  geom_boxplot(aes(x = logb0, y = su, color = factor(sb), group = factor(su)), outliers = FALSE)

# correlation between alpha and log_b0 ?
s.samples %>%
  as_draws() %>%
  spread_draws(alpha[su], logb0[su]) %>%
  median_qi() %>%
  ggplot(aes(y = alpha, x = logb0, xmin = logb0.lower, xmax = logb0.upper, ymin = alpha.lower, ymax = alpha.upper)) +
  geom_pointinterval(orientation = "vertical", point_size = 0.8, linewidth = 0.5) +
  geom_pointinterval(orientation = "horizontal", point_size = 0.8, linewidth = 0.5) 

# conditional L50 for mean age_sc (b2,b3 and b4 equals zero)
# s.samples %>%
#   as_draws() %>%
#   spread_draws(alpha[su], logb0[su]) %>%
#   mutate(L50_sc = -alpha/exp(logb0),
#          L50 = L50_sc * sd(df.sweel3$length) + mean(df.sweel3$length)) %>%
#   left_join(df.sweel3 %>% distinct(su,sb,marine,balance)) %>%
#   filter(balance == 1) %>%
#   ggplot() +
#   geom_density(aes(L50, color = factor(su)), show.legend = FALSE) +
#   theme_minimal() +
#   xlim(-1000,1000)
#   
# # conditional A50 for mean age_sc (b1,b3 and b4 equals zero)
# s.samples %>%
#   as_draws() %>%
#   spread_draws(alpha[su], b1) %>%
#   mutate(A50_sc = -alpha/b1,
#          A50 = A50_sc * sd(df.sweel3$age) + mean(df.sweel3$age),
#          ) %>%
#   left_join(df.sweel3 %>% distinct(su,sb,marine,balanced)) %>%
#   filter(balance == 1) %>%
#   ggplot() +
#   geom_density(aes(A50_sc, color = factor(su)), show.legend = FALSE) +
#   theme_minimal() +
#   #xlim(0,1500)
#   xlim(-15,15) 
# 
# # 50% transition surface
# s.samples %>%
#   as_draws() %>%
#   spread_draws(alpha[su], log_b0[su], b1) %>%
#   median_qi() %>%
#   expand_grid(age_sc = seq(min(df.sweel3$age_sc, na.rm = TRUE), max(df.sweel3$age_sc, na.rm = TRUE), length.out = 10)) %>%
#   mutate(length_sc = (-alpha - b1 * age_sc)/exp(log_b0)) %>%
#   ggplot() +
#   geom_line(aes(age_sc, length_sc, color = factor(su))) 
  
# sensitivity and specificity 
s.samples %>%
  as_draws() %>%
  #gather_draws(se_d,se_v,sp_d,sp_v) %>%
  gather_draws(se_v,sp_v) %>%
  ggplot() +
  geom_density(aes(x = .value, color = .variable)) +
  # stat_function(fun = dbeta, args = list(shape1 = 5, shape2 = 1), color = "black",linetype = "dashed") +
  # stat_function(fun = dbeta, args = list(shape1 = 10, shape2 = 1), color = "grey",linetype = "dashed") 
  xlim(0.9,1) 

# covariates 
s.samples %>%
  as_draws() %>%
  gather_draws(bA,bC,bE,bLa,bP,bH[sb]) %>%
  ggplot() +
  geom_density(aes(.value, color = factor(sb)), show.legend = FALSE) +
  facet_wrap(~.variable)
  
tmpcovs <- s.samples %>%
  as_draws() %>%
  gather_draws(p[ind_id],p_vis[ind_id],age_sc[ind_id],con_index_sc[ind_id],pec_index_sc[ind_id],eye_index_sc[ind_id]) %>%
  median_qi() %>%
  left_join(s.samples %>%
              as_draws() %>%
              spread_draws(p[ind_id]) %>% 
              median_qi(), by = "ind_id") 

tmpcovs %>%
  filter(!.variable %in% c("p","p_vis")) %>%
  left_join(tmpcovs %>%
              filter(!.variable %in% c("p","p_vis"))) %>%
  ggplot() +
  geom_point(aes(p, .value, color = .variable), alpha = 0.1) +
  facet_wrap(~.variable, scales = "free")
```


### Predicted / observed

```{r}
# var.lats <- data.silv %>%
#   mutate(m.lat = mean(ser_y), .by = ecoreg) %>%
#   distinct(er, ecoreg, m.lat) %>%
#   arrange(-m.lat)

dfp <- s.samples %>%
  as_draws() %>%
  spread_draws(mu_s[ind.id]) %>%
  median_qi() %>%
  rename(pred.sl = mu_s) %>%
  ungroup() %>%
  left_join(df.sweel %>%
              select(main_bas,length,lat,lon,sex,ind.id) %>%
              rename(obs.sl = length))
  #left_join(var.lats) %>%
  #mutate(ecoreg = fct_reorder(as.factor(ecoreg), -m.lat))

dfp %>%
  #filter(is.na(sex)) %>%
  ggplot() +
  geom_abline(slope = 1, color = "darkgrey") +
  geom_point(aes(x = obs.sl, y = exp(pred.sl), color = sex), shape = 4, alpha = 0.7, show.legend = FALSE) +
  facet_wrap(~main_bas) +
  theme_minimal() +
  labs(y="predicted", x="observed")
#ggsave(paste0(home, "/report/silv_po.png"), width = 17, height = 14, units = "cm")

```

### logistic p ~ length

Solve for z = 0 (i.e. when p = 0.5), which means that we want to get the observed length in the expression for $z$; 

$0 = alpha[su] + b_{L,su} * length_sc + b_{A,su} * age_sc + b_{H,su} * habitat + b_{1,su} * age_sc + b_{1,su} * age_sc$

Conditioned on mean age (age_sc = 0), this becomes:

$L50_{su} = -alpha[su] / b_{0,su}$

```{r}
# Adding covariate effects to the L50
# First make mean observed lat and inland/marine to each su
df_sucov <- df.sweel3 %>%
  summarise(lat_scmean = mean(lat_sc), 
            .by = su) %>%
  left_join(df.sweel3 %>% distinct(su,spat.unit,habitat,sb,balance,spat.unit) %>%
              mutate(coast = if_else(habitat == 1, "coast","inland"),
                     seaba = if_else(sb == 1, "Westcoast","Baltic Sea")))

df_coef <- s.samples %>%
  as_draws() %>%
  spread_draws(alpha[su], logb0[su]) %>%
  select(-.iteration,-.chain) %>%
  left_join(df_sucov %>% select(su,sb)) %>%
  left_join(s.samples %>%
              as_draws() %>%
              spread_draws(bA,bC,bH[sb],bE,bP,bLa), by = c("sb",".draw"))

df_data <- df.sweel3 %>%
  select(length,ind_id,su,sb,habitat,lat_sc) %>%
  left_join(s.samples %>%
              as_draws() %>%
              #spread_draws(age_sc[ind.id],logweight[ind.id]) %>%
              spread_draws(age_sc[ind_id],con_index_sc[ind_id],pec_index_sc[ind_id],eye_index_sc[ind_id]) %>%
              #spread_draws(age_sc[ind_id],pec_index[ind_id],eye_index[ind_id]) %>%
              median_qi()
              ) %>%
  # if le cren is not monitored but logweight is
  # cross_join(s.samples %>%
  #             as_draws() %>%
  #             spread_draws(ab[par]) %>%
  #             median_qi() %>%
  #             pivot_wider(values_from = c(.lower,.upper,ab), names_from = par)) %>%
  # mutate(logw_mu = ab_1 + ab_2*log(length),
  #        le_cren = exp(logweight)/exp(logw_mu)) %>%
  select(ind_id,con_index_sc,age_sc,su,habitat,lat_sc,pec_index_sc,eye_index_sc)

nsu <- length(unique(df.sweel3$su))
lengths <- seq(min(df.sweel3$length_sc), max(df.sweel3$length_sc), length.out = 100)
res <- vector("list", nsu)


for(j in 1:nsu){
  
  dat_j <- df_data %>% filter(su == j)
  coef_j <- df_coef %>% filter(su == j)
  
  res_j <- lapply(lengths, function(l){
    
    X_ind <- cbind(dat_j$age_sc, dat_j$con_index_sc, dat_j$eye_index_sc, dat_j$pec_index_sc, dat_j$habitat, dat_j$lat_sc)
    B_draws <- cbind(coef_j$bA, coef_j$bC, coef_j$bE, coef_j$bP, coef_j$bH, coef_j$bLa)
    z_unit <- coef_j$alpha + exp(coef_j$logb0)*l
  
    z_mat <- B_draws %*% t(X_ind) + as.vector(z_unit)  # n_draws x n_individuals
  
    p_mat      <- 1 / (1 + exp(-z_mat))
    p_marginal <- rowMeans(p_mat) # average over individuals: n_draws vector
  
    data.frame(
      su        = j,
      length_sc = l,
      p_mean    = mean(p_marginal),
      p_lo      = quantile(p_marginal, 0.025),
      p_up      = quantile(p_marginal, 0.975)
      )
    })
  
  res[[j]] <- bind_rows(res_j)
}

# for smaller datasets, dfl is both draws and data combined
# for(j in 1:nsu) {
#   tmp = dfl %>%
#     filter(su == j) %>%
#     expand_grid(length_sc = seq(min(df.sweel3$length_sc, na.rm = TRUE), max(df.sweel3$length_sc, na.rm = TRUE), by = 0.1)) %>%
#     p = 1 / (1 + exp(-(alpha + exp(logb0)*length_sc + b1*age_sc + b2*le_cren + b3*lat_scmean + b4*marine))) %>%
#     group_by(length_sc) %>%
#     median_qi(p) %>%
#     mutate(sup = j) %>%
#     bind_rows(t)
#   rm(tmp, tmp2)
#   gc()
#   }
dfs <- s.samples %>%
  as_draws() %>%
  spread_draws(silver[ind_id],con_index_sc[ind_id],age_sc[ind_id], pec_index_sc[ind_id],eye_index_sc[ind_id]) %>%
  median_qi() %>%
  left_join(df.sweel3 %>% distinct(ind_id,su,length,habitat,sb)) %>%
  #left_join(df_sucov) %>%
  ungroup() %>%
  mutate(coast = if_else(habitat == 1, "coast","inland"),
         sb = if_else(sb == 1, "Westcoast","Baltic Sea"))

# by each spatial unit
bind_rows(res) %>%
  mutate(length = length_sc*sd(df.sweel3$length, na.rm = TRUE) + mean(df.sweel3$length, na.rm = TRUE)) %>% 
  left_join(df_sucov %>% select(-sb)) %>%
  #filter(balance == "1") %>%
  ggplot() +
  geom_line(aes(length_sc, p_mean, color = seaba)) +
  geom_ribbon(aes(x = length_sc, y = p_mean, ymin = p_lo, ymax = p_up), alpha = 0.3, color = NA) +
  #, group = spat.unit), fill = factor(balance)), alpha = 0.5, color = NA) + 
#geom_point(data = df.sweel3, aes(length, y = silver_vis)) +
  geom_point(data = dfs %>% left_join(df.sweel3 %>% distinct(ind_id,length_sc,silver_vis)) %>% left_join(df_sucov %>% select(-sb)),
  #            #%>%  filter(balance == "1"), 9å
              aes(x = length_sc, y = silver_vis, color = factor(balance))) +
  facet_wrap(~su)

# by habitat and sea basin
bind_rows(res) %>%
  #left_join(df_sucov) %>%
  mutate(length = length_sc*sd(df.sweel3$length, na.rm = TRUE) + mean(df.sweel3$length, na.rm = TRUE)) %>% 
  left_join(df_sucov) %>%
  ggplot() +
  geom_line(aes(x = length, y = p_mean, group = lat_scmean, color = lat_scmean)) +
  geom_point(data = dfs %>% left_join(df.sweel3 %>% distinct(ind_id,length_sc,silver_vis)) %>% left_join(df_sucov %>% select(-sb)), aes(length, y = silver, group = lat_scmean, color = lat_scmean), show.legend = FALSE, alpha = 0.3, shape = 124, size = 1.5) +
  facet_wrap(coast~seaba) +
  scale_color_viridis_c()
ggsave(paste0(home, "/figures/L50_hab_basin.pdf"), width = 18, height = 13, units = "cm")

dfs %>%
  ggplot() +
  geom_density(aes(eye_index_sc, color = factor(silver))) +
  xlim(0.5,1.5)
```

### logistic p ~ age
Calculate for basin specific L50.

Solve for z = 0 (i.e. when p = 0.5), which means that we want to get the observed length in the expression for $z$; 

$0 = alpha[su] + b_{0,su} * length_sc + b_{1,su} * age_sc$

Conditioned on mean age (length = 0), this becomes:

$A50_{su} = -alpha[su] / b_{1,su}$

```{r}
s.samples %>%
  as_draws() %>%
  spread_draws(alpha[su], b1) %>%
  median_qi()

# using alpha, b0 which is per basin 
dfza <- s.samples %>%
  as_draws() %>%
  spread_draws(alpha[su], log_b0[su], b1) %>%
  median_qi() %>%
  expand_grid(age_sc = seq(min(df.sweel3$age_sc, na.rm = TRUE),max(df.sweel3$age_sc, na.rm = TRUE), by = 0.05)) %>%
  left_join(df.sweel3 %>% distinct(spat.unit,su)) %>%
  ungroup() %>%
  mutate(p_mean = 1 / (1 + exp(-(alpha + exp(log_b0)*0 + b1*age_sc))))

ssu <- sample(df.sweel3$su, 20, replace = FALSE)
dfza %>%
  filter(#sex == 0,
         su %in% ssu) %>%
  ggplot() +
  geom_line(aes(x = age_sc, y = p_mean, color = factor(su)), show.legend = FALSE) 
  #geom_point(data = df.sweel3 %>% filter(su %in% ssu), aes(age_sc, y = silver_dur, color = factor(spat.unit)),show.legend = FALSE, alpha = 0.3) 
  facet_wrap(~spat.unit) 

dfza %>%
  filter(sex == 0) %>%
  ggplot() +
  geom_line(aes(x = age_sc, y = p_mean, color = factor(su)), show.legend = FALSE)
```

```{r}
dfp <- silv.samples %>%
  spread_draws(s.mu[i]) %>%
  left_join(data.silv %>% 
              select(ecoreg,sex,fi_year) %>% 
              mutate(i = row_number()), 
            by = "i") %>%
  group_by(.draw, ecoreg, sex, fi_year) %>%
  summarise(pred.sl = median(s.mu), .groups = "drop") %>%
  left_join(var.lats) %>%
  mutate(ecoreg = fct_reorder(as.factor(ecoreg), m.lat),
         sex = if_else(sex == 0, "female", "male")) %>%
  mutate(m_pred = median(pred.sl), .by = c(sex))

n.labs <- data.silv %>%
  summarise(n = n(), .by = c(ecoreg, sex)) %>%
  mutate(sex = if_else(sex==1, "female", "male"))

dfp %>% summarise(mexp = exp(median(pred.sl)), .by= sex)
dfp %>% ungroup() %>% summarise(exp(median(pred.sl)), .by= sex)

dfp %>%
  ggplot() +
  geom_boxplot(aes(x = exp(pred.sl), y = ecoreg), outliers = FALSE, show.legend = FALSE) +
  geom_vline(aes(xintercept = exp(m_pred)), color = "darkgrey", size = 0.5) +
  facet_wrap(~sex, scales = "free_x") +
  theme_minimal() +
  labs(x = "Length at silvering [mm]", y = "ecoregion")
ggsave(paste0(home, "/report/silv_ERl.png"), width = 17, height = 14, units = "cm")

```

### Length over space

```{r, cache = FALSE}
# L50 by points
dfs <- s.samples %>%
  as_draws() %>%
  spread_draws(alpha[su], log_b0[su], sep = ",") %>%
  median_qi() %>%
  mutate(L50 = -alpha/exp(log_b0)) %>%
  # gather_draws(alpha[mb], bs[er], bt, sep = ",") %>%
  left_join(df.sweel3 %>% distinct(spat.unit,su,lat,lon))# %>%

# by catchment
dfsca <- ca_sf %>%
  mutate(spat.unit = as.character(main_bas)) %>%
  #mutate(main_bas = as.factor(main_bas)) %>%
  left_join(dfs %>% distinct(su,L50,spat.unit)) %>%
  filter(!is.na(mb)) #%>%
  mutate(pred.length = exp(.value))
  
plot_map_Nord +
  geom_sf(data = dfs, aes(fill = L50), color = NA,  alpha = 0.5) +
  scale_fill_viridis_c() +
  geom_point(data = df.sweel2 %>% drop_na(lat,lon) %>% 
               add_utm_columns(ll_names = c("lon", "lat"), utm_crs = 32633), aes(X*1000, Y*1000), #color = , 
             size = 0.1) 
  
# # ungroup() %>%
#   # mutate(female = alpha + bt*mt,
#   #        male = alpha + bs + bt*mt,
#   #        year = yr + 1997) %>%
#   pivot_longer(cols = c("female","male"), names_to = "sex", values_to = "length") %>%
#   mutate(length = exp(length)) %>%
#   mutate(m_length = median(length), .by = c(year,sex)) %>%
#   left_join(var.lats) %>%
#   mutate(ecoreg = fct_reorder(ecoreg, -m.lat)) 
# 
# dfp %>%
#   filter(sex == "female") %>%
#   ggplot() +
#   geom_boxplot(aes(y = length, x = year, group = year), color = "#F8766D", outliers = FALSE) +
#   geom_line(aes(y = m_length, x = year), color = "#F8766D", alpha = 0.4) +
#   facet_wrap(~ecoreg, scales = "free_y") +
#   theme_minimal() +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1, size = rel(.75))) +
#   labs(x = "year")
# ggsave(paste0(home, "/report/silv_lf.png"), width = 17, height = 14, units = "cm")
# 
# dfp %>%
#   filter(sex == "male") %>%
#   ggplot() +
#   geom_boxplot(aes(y = length, x = year, group = year), color = "#00BFC4", outliers = FALSE) +
#   geom_line(aes(y = m_length, x = year), color = "#00BFC4", alpha = 0.4) +
#   facet_wrap(~ecoreg, scales = "free_y") +
#   theme_minimal() +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1, size = rel(.75))) +
#   labs(x = "year")
# ggsave(paste0(home, "/report/silv_lm.png"), width = 17, height = 14, units = "cm")  

```

### Length distribution per su

```{r}
pis <- s.samples %>%
  as_draws() %>%
  #spread_draws(p[ind.id], length_sc[ind.id], age_sc[ind.id], weight_sc[ind.id], silver[ind.id]) %>%
  spread_draws(p[ind_id], silver[ind_id], silver_vis[ind_id]) %>%
  median_qi() %>%
  left_join(df.sweel3 %>% select(su,spat.unit,ind_id,sb,balance,marine,sea_basin,length), by = "ind_id") 

# probability vs silver
pis %>%
  ggplot() +
  geom_jitter(aes(y = silver, x = p, color = length), alpha = 0.7, position = "jitter", size = 0.3) 

# Weighting by p has very little effect on the density distribution:
pis %>%
  #mutate(length = length_sc * sd(df.sweel3$length, na.rm = TRUE) + mean(df.sweel3$length, na.rm = TRUE)) %>%
  filter(silver == 1,
         balance == 1) %>%
  ggplot() + 
  geom_density_ridges(aes(x = length, y = sb, weight = p, fill = factor(sb)), color = NA, alpha = 0.5) +
  geom_density_ridges(aes(x = length, y = sb, color = factor(balance)), fill = NA) +
  labs(caption = "filled is weigthed by p")

# density by balanced sunits and sb 
p3 <- pis %>%
  mutate(length = length_sc * sd(df.sweel3$length, na.rm = TRUE) + mean(df.sweel3$length, na.rm = TRUE),
         sb = if_else(sb == 1, "Westcoast", "Baltic Sea")
         ) %>%
  filter(silver == 1,
         balance == 1
         ) %>%
  ggplot() + 
  geom_density_ridges(aes(x = length, y = spat.unit, weight = p, fill = factor(sb)),alpha = 0.8, color = NA)

pis %>%
#  left_join(df.sweel3 %>% select(ind_id,marine), by = "ind_id") %>%
  # mutate(habitat = if_else(marine == 1, "coast","inland"),
  #        sb = if_else(sb == 1, "Westcoast","Baltic Sea"),
  #        #length = length_sc * sd(df.sweel3$length, na.rm = TRUE) + mean(df.sweel3$length, na.rm = TRUE),
        # ) %>%
  filter(silver == 1) %>%
  ggplot(aes(x = length, y = sb, weight = p, fill = marine)) + 
  geom_density_ridges()

p3 / p4 + plot_layout(ncol = 1, guides = "collect")

plot_map_Nord +
  geom_point(data = df.sweel3 %>% 
               filter(su %in% c(6,27)) %>%
               filter(balance == 1) %>% add_utm_columns(ll_names = c("lon", "lat"), utm_crs = 32633), aes(X*1000, Y*1000, color = sea_basin)) +
  geom_sf(data = ca_sf2, aes(fill = factor(main_catch)),  alpha = 0.3)

# 6, 27,59
```

### Posterior predictive checks
```{r}
# visual classifier
# compare means by spatial unit
tmps <- as_draws(s.samples) %>%
  #subset_draws(variable = c("visual_rep")) %>%
  spread_draws(visual_rep[ind_id],silver_rep[ind_id],silver[ind_id]) %>%
  median_qi() %>%
  left_join(df.sweel3 %>% select(su,spat.unit,ind_id,silver_vis,durif_stage), by = "ind_id") 

tmps %>%
  summarise(mean_est_vis = mean(visual_rep, na.rm = TRUE), .by = spat.unit) %>%
  left_join(df.sweel3 %>% summarise(obs_vis = mean(silver_vis, na.rm = TRUE), .by = spat.unit), by = "spat.unit") %>%
  left_join(df_sucov) %>%
  ggplot() +
  geom_point(aes(obs_vis,mean_est_vis)) +
  geom_abline() 

tmps %>%
  left_join(df.sweel3 %>% select(durif_stage,ind_id,silver_vis,length_sc)) %>%
  summarise(n = n(), ml = mean(length_sc), .by = c(durif_stage,visual_rep)) %>%
  arrange(-n) 

tmps %>%
  left_join(df.sweel3 %>% select(durif_stage,ind_id,silver_vis,length)) %>%
  summarise(n = n(), ml = mean(length), .by = c(silver_vis,silver_rep,visual_rep)) %>%
  arrange(-n) 

left_join(df_sucov) %>%
  ggplot() +
  geom_point(aes(durif_stage,visual_rep)) 

# plot densities of 0/1 
tmps %>%
  ggplot() +
  geom_density(aes(visual_rep), color = "red") +
  geom_density(data = df.sweel3, aes(silver_vis)) +
  labs(title = "red is model") 

tmps %>%
  summarise(mean_est_vis = sum(visual_rep, na.rm = TRUE), .by = spat.unit) %>%
  left_join(df.sweel3 %>% summarise(obs_vis = sum(silver_vis, na.rm = TRUE), .by = spat.unit), by = "spat.unit") %>%
  left_join(df_sucov) %>%
  ggplot() +
  geom_point(aes(obs_vis,mean_est_vis, color = factor(balance))) +
  geom_abline()

# for the balancing by inroducing NAs model
tmps2 <- df.sweel3 %>%
  mutate(pre_silver_vis = case_when(visual_stage == "Blankål" ~ 1,
                                visual_stage == "Halvblank ål" ~ NA,
                                visual_stage %in% c("Gulål","Yngel","Juvenil") ~ 0,
                                .default = NA)) %>%
  select(silver_vis,pre_silver_vis,ind_id) %>%
  left_join(as_draws(s.samples) %>%
              spread_draws(silver[ind_id]) %>%
              median_qi()) 
tmps2 %>%
  # select those that were set to NA before model
  filter(is.na(pre_silver_vis) == !is.na(silver_vis)) %>%
  summarise(n(), .by = c(silver,pre_silver_vis)) 
9238/(5739+9283)
2733/(4812+2733)


# Durif classifier
# tmpd <- as_draws(s.samples) %>%
#   #subset_draws(variable = c("visual_rep")) %>%
#   spread_draws(measured_rep[ind.id]) %>%
#   median_qi() %>%
#   left_join(df.sweel3 %>% select(su,spat.unit,ind.id), by = "ind.id") 
# 
# tmpd %>%
#   summarise(mean_est_dur = mean(measured_rep, na.rm = TRUE), .by = spat.unit) %>%
#   left_join(df.sweel3 %>% summarise(obs_dur = mean(silver_dur, na.rm = TRUE), .by = spat.unit), by = "spat.unit") 
# 
# tmpd %>%
#   ggplot() +
#   geom_density(aes(measured_rep), color = "red") +
#   geom_density(data = df.sweel3, aes(silver_vis)) +
#   labs(title = "red is model")
#   ggplot() +
#   geom_density(aes(measured_rep)) +
#   geom_density(data = df.sweel3, aes(silver_dur))

# Test
# Disagreement rate between classifiers — the model's ability to reproduce the observed rate of V≠M is a direct check on Se/Sp estimation
#Prevalence by population — check that simulated stage prevalence matches observed classifier prevalence within each population mb
#Prevalence along the length gradient — bin individuals by length and compare observed vs simulated classifier rates, which checks the logistic curve f
```

### The durifs

```{r}
tmp3s <- s.samples %>%
  as_draws() %>%
  spread_draws(silver[ind_id]) %>%
  median_qi() %>%
  left_join(df.sweel3 %>% select(visual_stage,durif_stage,ind_id,su,balance,length), by = "ind_id")

tmp3s %>%
  filter(durif_stage %in% c("FII","FIII","FIV","FV")) %>%
  summarise(n = n(), .by = c(silver, durif_stage, visual_stage)) %>%
  arrange(-silver, -n) %>%
  rename(estimated = silver)


```
### Compare visual and durif with classifier

```{r}
tmp4 <- s.samples %>%
  as_draws() %>%
  spread_draws(silver[ind_id],p[ind_id],p_vis[ind_id]) 

tmp4b <- tmp4 %>%
  group_by(ind_id) %>%
  median_qi() %>%
  left_join(df.sweel3 %>% select(silver_vis,visual_stage,durif_stage,ind_id,su,length,balance), by = "ind_id")

tmp4b %>%
  ggplot() +
  geom_point(aes(length,p), alpha =0.1) +
  facet_wrap(~su)

tmp4b %>%
  ggplot(aes(x=length,y=p)) +
  geom_point( alpha =0.1) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 3), show.legend = FALSE) +
  facet_wrap(~silver_vis)
# tmp4b <- tmp4 %>%
#   filter(silver == 1,
#          balance == 1) %>%
#   summarise(mean = mean(length),
#             sd = sd(length), .by = su) %>%
#   ungroup() %>%
#   cbind(tmp4 %>%
#           filter(visual_stage == "Blankål",
#                  balance == 1) %>%
#           summarise(meanv = mean(length),
#                     sdv = sd(length), .by = su) %>%
#           ungroup() %>%
#           select(-su))

tmp4c <- tmp4b %>%
  filter(p > 0.75,
         balance == 1) %>%
  summarise(mean = mean(length),
            sd = sd(length), .by = su) %>%
  ungroup() %>%
  mutate(source = "model") %>%
  rbind(tmp4b %>%
          filter(visual_stage == "Blankål",
                 balance == 1) %>%
          summarise(mean = mean(length),
                    sd = sd(length), .by = su) %>%
          ungroup() %>%
          mutate(source = "visual")) %>%
  rbind(tmp4b %>%
          filter(durif_stage %in% c("FIII","FIV","FV"),
                 balance == 1) %>%
          summarise(mean = mean(length),
                    sd = sd(length), .by = su) %>%
          ungroup() %>%
          mutate(source = "Durif"))

tmp4c %>%
  expand_grid(length = seq(400, 1400, length.out = 300)) %>%
  mutate(y = dnorm(length, mean = mean, sd = sd)) %>%
#  filter(su %in% c(15)) %>%
  ggplot() +
  #geom_area(aes(length, y, fill = source), position = "identity", alpha = 0.4) +
  geom_line(aes(length, y,  color = source)) +
  facet_wrap(~su) +
  geom_vline(xintercept = mean(df.sweel3$length), color = "darkgrey") 
ggsave(paste0(home, "/figures/compare_classif.pdf"), width = 18, height = 13, units = "cm")

# tmp5 %>%
#   filter(silver == 1) %>%
#   ggplot() +
#   geom_density(aes(length)) +
#   geom_vline(data = tmp5 %>% filter(silver == 1) %>% mutate(m = mean(length), .by = su), aes(xintercept = m)) +
#   geom_density(data = tmp5 %>% filter(visual_stage == "Blankål"), aes(length), color = "red") +
#   geom_vline(data = tmp5 %>% filter(visual_stage == "Blankål") %>% mutate(m = mean(length), .by = su), aes(xintercept = m), color = "red") +
#   geom_density(data = tmp5 %>% filter(durif_stage %in% c("FIII","FIV","FV")), aes(length), color = "blue") +
#   geom_vline(data = tmp5 %>% filter(durif_stage %in% c("FIII","FIV","FV")) %>% mutate(m = mean(length), .by = su), aes(xintercept = m), color = "blue") +
#   facet_wrap(~su)

```

# test distributions
```{r}
mu_a_gl ~ dnorm(msa, sda)
sd_a_gl ~ dgamma(shape = 1.5, rate = 1.5 / sda)
#mu_b_gl ~ dnorm(log(mb0), sd = log(sdb0))
mu_b_gl ~ dnorm(log(mb0), sd = 0.5)
#sd_b_gl ~ dgamma(shape = 1.5, rate = 1.5 / sdb0) # to weak 
sd_b_gl ~ dgamma(shape = 2, rate = 4 / sdb0) 

sd_a_su ~ dgamma(shape = 2, rate = 2cx)   # sd for su-level
  

ggplot() +
  #stat_function(fun = dgamma, 
  #            args = list(shape = 1000, rate = 2000)) +
  stat_function(fun = dgamma, 
              args = list(shape = 2,rate=4), color = "blue") + xlim(0,5)  
ggplot() +
  #stat_function(fun = dgamma, 
  #            args = list(shape = 1000, rate = 2000)) +
  stat_function(fun = dbeta, 
              args = list(5,2), color = "blue") 
  stat_function(fun = dexp, 
              args = list(7), color = "blue") +
  stat_function(fun = dexp, 
              args = list(3)) +
  xlim(0,5)
```

```{r}
tmpab <- 
  subset_draws(as_draws(s.samples), variable = c("ab")) %>%
  summarise_draws(c("mean","sd")) %>%
  mutate(variable = case_when(variable == "ab[1]" ~ "a",
                              variable == "ab[2]" ~ "b")) %>%
  pivot_wider(names_from = variable, values_from = c(mean,sd))
                           
df.sweel3 %>%
  select(length,weight,su) %>%
  drop_na(length,weight) %>%
  bind_cols(tmpab) %>%
  mutate(logweight_mu = mean_a + mean_b*log(length),
         le_cren = weight / exp(logweight_mu)) %>%
  filter(le_cren < 0.001) %>%
  ggplot() +
  #geom_point(aes(length, weight))
  geom_density(aes(le_cren
                  ))
```

