---
title: "Swedish eel silvering model summary"
subtitle: "SWEELHT"
author: "Viktor Thunell" 
date: "`r format(Sys.time(), '%d %B, %Y')`"
format: 
  html:
    code-fold: true
    code-summary: "Show code"
    page-layout: full
    embed-resources: true
    toc: true
knitr: 
  opts_chunk:
    fig.align: center
    out-width: 80%
editor: source
execute: 
  echo: true
  eval: true
  cache: true
---

# Latent classs logistic regression for eel migration probability

## Load libraries

```{r}
#| message: false
#| warning: false
#| cache: false

# Load libraries, install if needed
pkgs <- c("tidyverse", "tidylog", "devtools","viridis","sf","patchwork","posterior","ggridges", "tidybayes","bayesplot","nimble","here","reshape2","sdmTMB")

if(length(setdiff(pkgs,rownames(installed.packages()))) > 0){
    install.packages(setdiff(pkgs, rownames(installed.packages())), dependencies = T)
  }

invisible(lapply(pkgs, library, character.only = T))

devtools::source_url("https://raw.githubusercontent.com/VThunell/diasp-lht/main/R/functions/map-plot.R")
options(ggplot2.continuous.colour = "viridis")
theme_set(theme_minimal()) 

# Set path
home <- here::here()
```

## Read data

```{r}
#| message: false
#| warning: false
#| cache: false

## Read data
sweels <- readRDS(file = paste0(home, "/data/sweel_data_2026-06-24.RData"))
```

## Silvering logistic regression model

### Define data

```{r}
# df.sweel2 <- sweels %>%
#   drop_na(length) %>%
#   filter(visual_stage %in% c("Blankål","Halvblank ål","Gulål"),
#          country == "SE") %>%
#   mutate(silver = case_when(visual_stage == "Blankål" ~ 1,
#                             visual_stage == "Halvblank ål" ~ NA,
#                             visual_stage == "Gulål" ~ 0,
#                             .default = NA),
#          silver = as.integer(silver),
#          length_sc = (length - mean(length)) / sd(length),
#          mb = as.integer(factor(main_bas)),
#          sex = as.integer(if_else(sex == "female", 0, 1)),
#          ind.id = row_number())

df.sweel3 <- sweels %>%
  mutate(weight = if_else(weight == 0, NA, weight)) %>%
  # remove rows where used variables have at least two non NAs
  filter(rowSums(!is.na(select(., visual_stage, durif_stage, weight, length, age))) >= 2,
         !(is.na(visual_stage) & is.na(durif_stage)),
         sex == "female") %>% 
  mutate(silver_vis = case_when(visual_stage == "Blankål" ~ 1,
                                visual_stage == "Halvblank ål" ~ NA,
                                # OR: visual_stage == "Halvblank ål" ~ NA,
                                visual_stage %in% c("Gulål","Yngel","Juvenil") ~ 0,
                                .default = NA),
         silver_dur = case_when(durif_stage %in% c("FIV","FV") ~ 1,
                                durif_stage %in% c("FIII") ~ NA,
                                durif_stage %in% c("I","FII","MII") ~ 0,
                                .default = NA),
         length_sc = (length - mean(length, na.rm = TRUE)) / sd(length, na.rm = TRUE),
         age_sc = (age - mean(age, na.rm = TRUE)) / sd(age, na.rm = TRUE),
         weight_sc = (weight - mean(weight, na.rm = TRUE)) / sd(weight, na.rm = TRUE),
         logweight = log(weight),
         # create dummy su for NA basins
         spat.unit = if_else(is.na(spat.unit),str_c("dum_",sea_basin),spat.unit),
         su = as.integer(factor(spat.unit)),
         sb = case_when(sea_basin == "VH" ~ 1,
                        !sea_basin == "VH" ~ 2),
         #sb = as.integer(factor(sea_basin)),
         habitat = if_else(marine == "yes", 1,0),
         lat_sc = (lat - mean(lat, na.rm = TRUE)) / sd(lat, na.rm = TRUE),
         ind.id = row_number()) 

# which sites are unbalanced?
su.bal <- df.sweel3 %>%
  rowwise() %>%
  mutate(silver = max(c(silver_vis,silver_dur), na.rm = TRUE)) %>%
  ungroup() %>%
  summarise(n = n(), .by = c(silver,spat.unit)) %>%
  pivot_wider(id_cols = spat.unit, names_from = silver, values_from = n) %>%
  mutate(ratio = `1`/`0`,
         balance = if_else(ratio < 0.1 | ratio > 10 | is.na(ratio), 0, 1)) %>%
  select(spat.unit,balance) %>%
  # the dummy su:s (Östersjön and Västerhavet) should only contribute to the sb level and thus be unbalanced
  mutate(balance =  if_else(str_detect(spat.unit, "dum_"), 0, balance))

# join in balance
df.sweel3 <- df.sweel3 %>%
  left_join(su.bal)

df.sweel3 %>%
  distinct(silver_vis,silver_dur)

df.sweel3 %>%
  summarise(n = n(), .by = c(silver_vis,silver_dur)) %>%
  arrange(-silver_vis,-silver_dur)

# the sb:s have data so we can get sb level nested hierarchy
df.sweel3 %>%
  filter(!(is.na(silver_vis) & is.na(silver_dur))) %>%
  rowwise() %>%
  mutate(silver = max(c(silver_vis,silver_dur), na.rm = TRUE)) %>%
  ungroup() %>%
  summarise(n = n(), .by = c(silver,sea_basin,sb)) %>%
  pivot_wider(id_cols = c(sea_basin,sb), names_from = silver, values_from = n)

```

### Plot silvering data
```{r}
# length at age by stage and sex
df.sweel3 %>%
  ggplot(aes(age, length, color = factor(sex))) +
  geom_point(size = 0.1, alpha = 0.3) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), show.legend = FALSE) 
  facet_wrap(~visual_stage)
  
df.sweel3 %>%
  ggplot(aes(age, length, color = factor(sex))) +
  geom_point(size = 0.1, alpha = 0.3) +
  facet_wrap(~durif_stage) 

# relationship between age and silver
df.sweel3 %>%
  mutate(silv = silver_vis * silver_dur,
         silv = if_else(is.na(silv), 0.5, silv)) %>%
  ggplot(aes(silv, age)) +
  geom_point(size = 0.1, alpha = 0.3) +
  geom_smooth(method = "lm", se = TRUE) +
  theme_minimal()
  
# relationship between length and silver
df.sweel3 %>%
  mutate(silv = silver_vis * silver_dur,
         silv = if_else(is.na(silv), 0.5, silv)) %>%
  ggplot(aes(silv, length)) +
  geom_point(size = 0.1, alpha = 0.3) +
  geom_smooth(method = "lm", se = TRUE) +
  theme_minimal()

# relationship between length and silver
df.sweel3 %>%
  mutate(silv = silver_vis * silver_dur,
         silv = if_else(is.na(silv), 0.5, silv)) %>%
  ggplot() +
  geom_histogram(aes(length, fill = factor(tu)), alpha = 0.5) +
  facet_wrap(~silv)

# length of the silvers (where one of the classifiers says it is silver)
df.sweel3 %>%
  mutate(silver_dur = if_else(is.na(silver_dur),-1,silver_dur),
         silver_vis = if_else(is.na(silver_vis),-1,silver_vis),
         silv = silver_dur+silver_vis) %>%
  filter(!silv < 0) %>%
  ggplot(aes(length, factor(su), color = db)) +
  geom_boxplot(outlier.color = "grey",
               outlier.size = 0.5) 

# LHTs of Durif staged eels
df.sweel3 %>%
  filter(!is.na(durif_stage)) %>%
  ggplot(aes(length, factor(su), color = durif_stage)) +
  geom_boxplot(outlier.color = "grey",
               outlier.size = 0.5) 
df.sweel3 %>%
  filter(!is.na(durif_stage)) %>%
  ggplot(aes(age, factor(su), color = durif_stage)) +
  geom_boxplot(outlier.color = "grey",
               outlier.size = 0.5) 
df.sweel3 %>%
  filter(!is.na(durif_stage)) %>%
  ggplot(aes(weight1, factor(su), color = durif_stage)) +
  geom_boxplot(outlier.color = "grey",
               outlier.size = 0.5) 

# length of visual staged eels
df.sweel3 %>%
  filter(!is.na(visual_stage)) %>%
  ggplot(aes(length, factor(su), color = visual_stage)) +
  geom_boxplot(outlier.color = "grey",
               outlier.size = 0.5) 

# length-weight
df.sweel3 %>%
  drop_na(length,weight1) %>%
  ggplot(aes(length, weight1,color = factor(silver_dur))) +
  geom_point(size = 0.1, alpha = 0.3) 

# numbers of inds per clssifier    
df.sweel3 %>%
  summarise(n = n(), .by = c(visual_stage,durif_stage)) %>%
  arrange(visual_stage,-n)

df.sweel3 %>%
  filter(!(is.na(silver_vis) & is.na(silver_dur))) %>%
  rowwise() %>%
  mutate(silver = max(c(silver_vis,silver_dur), na.rm = TRUE)) %>%
  ungroup() %>%
  summarise(n = n(), .by = c(silver,spat.unit)) %>%
  pivot_wider(id_cols = spat.unit, names_from = silver, values_from = n) %>%
  arrange(-`1`,spat.unit)

df.sweel3 %>%
  filter(!(is.na(silver_vis) & is.na(silver_dur))) %>%
  rowwise() %>%
  mutate(silver = max(c(silver_vis,silver_dur), na.rm = TRUE)) %>%
  ungroup() %>%
  summarise(n = n(), .by = c(silver,sea_basin)) %>%
  pivot_wider(id_cols = sea_basin, names_from = silver, values_from = n) %>%
  arrange(-`1`,spat.unit)
```
### n over space and time 

```{r}
df.sweel3 %>%
  summarise(count = n(), .by = c(su))

df.sweel3 %>%
  summarise(count = n(), .by = c(year, su)) %>%
  ggplot(aes(year, factor(su), fill = count)) +
  geom_tile(color = "gray30") +
  #geom_vline(xintercept = 1998, color = "darkgrey") +
  scale_fill_viridis_c() +
  labs(y = "", x = "year") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Source models / Load samples
```{r, cache = FALSE}
# Load samples
sweelsilv.samples <- readRDS(paste0(home,"/data/samples/sweelsilv.samples_e2026-06-24.RData"))
sweelsilv.samples <- readRDS(paste0(home,"/data/samples/sweelsilv.samples_d2026-06-23.RData"))
sweelsilv.samples$WAIC
sweelsilv.samples_d$WAIC
# for simple 33029.01 - 20260507
# for spatial alpha 13956.79 - 20260507
# for spatial and hierachical alpha 13875.92 - 20260508
# for spatial b0 and hierachical spatial alpha 13570.31 - 20260508
# for spatial and hierachical alpha and b0 13506.59 - 20260511
# for spatial and hierachical alpha and b0 21417.19 - 20260511 (this adds data so incomparable to previous models)
# for spatial and hierachical alpha and b0 with two classifiers and se/sp: 8411.084
# 225708 # new data and model after meeting 4/6 (model 4, no condition)
#  234143.4 20260615 and then 231746.8 for the 4c model with 
# 232228.8 4c
# 231643.5 4e
# 231671 4e 22june
# 44273.07 4d 22june
s.samples <- sweelsilv.samples$samples
# # Or source model code and mcmc (load libs and data)
# t <- Sys.time()
# source(file = paste0(home,"/models_eel/model_eel_silvering.R"))
# Sys.time() - t # approx 4.5 hours with 10000 hmc samples
# WAIC based on combined log prob of silver_vis and _dur; d: 25569.66 (22946.08), e:26438.2
ll_vis <- s.samples[, grep("logProb_silver_vis", colnames(s.samples))]
ll_dur <- s.samples[, grep("logProb_silver_dur", colnames(s.samples))]
ll_total <- ll_vis + ll_dur  # n_draws x n_individuals

# Compute WAIC manually
lppd    <- sum(log(colMeans(exp(ll_total))))
p_waic  <- sum(apply(ll_total, 2, var))
-2 * (lppd - p_waic)

```

### Traces, ess, and ‘potential scale reduction factor’

Variables "s.sig","Ustar","U","mu.a.gl","sd.a.gl","alpha","bs","bt")
```{r, cache = FALSE}
variable.names(as_draws(s.samples)) %>%
  str_remove("\\[.*\\]") %>% unique()

node.sub <- subset_draws(as_draws(s.samples), variable = c("alpha_sb","logb0_sb","mu_a_gl","mu_b_gl"))
node.sub[] %>% mcmc_trace() 

node.sub <- subset_draws(as_draws(s.samples), variable = c("alpha"))
node.sub[,,sample(1:dim(node.sub)[3], 16)] %>% mcmc_trace() # two chain
node.sub[,sample(1:dim(node.sub)[2], 16)] %>% mcmc_trace() # one chain
node.sub %>% summarise_draws(c("mean","median","sd","rhat","ess_bulk","ess_tail")) 

node.sub <- subset_draws(as_draws(s.samples), variable = c("logb0"))
node.sub[,,sample(1:dim(node.sub)[3], 16)] %>% mcmc_trace() # one chain
node.sub %>% summarise_draws(c("mean","median","sd","rhat","ess_bulk","ess_tail"))

node.sub <- subset_draws(as_draws(s.samples), variable = c("sp_v","sp_d","se_v","se_d","b1","b2","b3","b4","ab","w_nu","w_sd"))
node.sub[] %>% mcmc_trace() # one chain
node.sub %>% summarise_draws(c("mean","median","sd","rhat","ess_bulk","ess_tail"))

node.sub <- subset_draws(as_draws(s.samples), variable = "p")
#node.sub[,,sample(1:dim(node.sub)[3], 16)] %>% mcmc_trace()
node.sub[,sample(1:dim(node.sub)[2], 36)] %>% mcmc_trace()
#node.sub %>% mcmc_trace() # one chain
node.sub[,sample(1:dim(node.sub)[2], 600)] %>% summarise_draws(c("median","sd","rhat","ess_bulk","ess_tail")) %>% arrange(-rhat)

node.sub <- subset_draws(as_draws(s.samples), variable = "silver_rep")
node.sub <- subset_draws(as_draws(s.samples), variable = "measured_rep")
node.sub[,sample(1:dim(node.sub)[2], 36)] %>% mcmc_trace()
#node.sub %>% mcmc_trace() # one chain
node.sub[,sample(1:dim(node.sub)[2], 600)] %>% summarise_draws(c("median","sd","rhat","ess_bulk","ess_tail")) %>% arrange(-rhat)

```

### Silvering variables

```{r}
# alpha_sb/beta0_sb (sea basin unit), baltic vs west coast
s.samples %>%
  as_draws() %>%
  spread_draws(alpha_sb[sb]) %>%
  ggplot() +
  geom_density(aes(x = alpha_sb, color = factor(sb))) +

s.samples %>%
  as_draws() %>%
  spread_draws(logb0_sb[sb]) %>%
  ggplot() +
  geom_density(aes(x = logb0_sb, color = factor(sb))) 

# alpha an b0 by su
s.samples %>%
  as_draws() %>%
  spread_draws(alpha[su]) %>%
  left_join(df.sweel3 %>% distinct(su,sb,marine)) %>%
  ungroup() %>%
  ggplot() +
  geom_boxplot(aes(x = alpha, y = su, color = factor(sb), group = factor(su)), outliers = FALSE) +

s.samples %>%
  as_draws() %>%
  spread_draws(logb0[su]) %>%
  left_join(df.sweel3 %>% distinct(su,sb,marine)) %>%
  ggplot() +
  geom_boxplot(aes(x = logb0, y = su, color = factor(sb), group = factor(su)), outliers = FALSE)

# correlation between alpha and log_b0 ?
s.samples %>%
  as_draws() %>%
  spread_draws(alpha[su], logb0[su]) %>%
  median_qi() %>%
  ggplot(aes(y = alpha, x = logb0, xmin = logb0.lower, xmax = logb0.upper, ymin = alpha.lower, ymax = alpha.upper)) +
  geom_pointinterval(orientation = "vertical", point_size = 0.8, linewidth = 0.5) +
  geom_pointinterval(orientation = "horizontal", point_size = 0.8, linewidth = 0.5) 

# conditional L50 for mean age_sc (b2,b3 and b4 equals zero)
# s.samples %>%
#   as_draws() %>%
#   spread_draws(alpha[su], logb0[su]) %>%
#   mutate(L50_sc = -alpha/exp(logb0),
#          L50 = L50_sc * sd(df.sweel3$length) + mean(df.sweel3$length)) %>%
#   left_join(df.sweel3 %>% distinct(su,sb,marine,balance)) %>%
#   filter(balance == 1) %>%
#   ggplot() +
#   geom_density(aes(L50, color = factor(su)), show.legend = FALSE) +
#   theme_minimal() +
#   xlim(-1000,1000)
#   
# # conditional A50 for mean age_sc (b1,b3 and b4 equals zero)
# s.samples %>%
#   as_draws() %>%
#   spread_draws(alpha[su], b1) %>%
#   mutate(A50_sc = -alpha/b1,
#          A50 = A50_sc * sd(df.sweel3$age) + mean(df.sweel3$age),
#          ) %>%
#   left_join(df.sweel3 %>% distinct(su,sb,marine,balanced)) %>%
#   filter(balance == 1) %>%
#   ggplot() +
#   geom_density(aes(A50_sc, color = factor(su)), show.legend = FALSE) +
#   theme_minimal() +
#   #xlim(0,1500)
#   xlim(-15,15) 
# 
# # 50% transition surface
# s.samples %>%
#   as_draws() %>%
#   spread_draws(alpha[su], log_b0[su], b1) %>%
#   median_qi() %>%
#   expand_grid(age_sc = seq(min(df.sweel3$age_sc, na.rm = TRUE), max(df.sweel3$age_sc, na.rm = TRUE), length.out = 10)) %>%
#   mutate(length_sc = (-alpha - b1 * age_sc)/exp(log_b0)) %>%
#   ggplot() +
#   geom_line(aes(age_sc, length_sc, color = factor(su))) 
  
# sensitivity and specificity 
# s.samples %>%
#   as_draws() %>%
#   gather_draws(se_d,se_v,sp_d,sp_v) %>%
  ggplot() +
  #geom_density(aes(x = .value, color = .variable)) 
  stat_function(fun = dbeta, args = list(shape1 = 5, shape2 = 1), color = "black",linetype = "dashed") +
  stat_function(fun = dbeta, args = list(shape1 = 10, shape2 = 1), color = "grey",linetype = "dashed") 
  xlim(0.9,1) 

# covariates 
s.samples %>%
  as_draws() %>%
  gather_draws(b1,b2,b3,b4[sb]) %>%
  ggplot() +
  geom_density(aes(.value, color = factor(sb)), show.legend = FALSE) +
  facet_wrap(~.variable)
  

```


### Predicted / observed

```{r}
# var.lats <- data.silv %>%
#   mutate(m.lat = mean(ser_y), .by = ecoreg) %>%
#   distinct(er, ecoreg, m.lat) %>%
#   arrange(-m.lat)

dfp <- s.samples %>%
  as_draws() %>%
  spread_draws(mu_s[ind.id]) %>%
  median_qi() %>%
  rename(pred.sl = mu_s) %>%
  ungroup() %>%
  left_join(df.sweel %>%
              select(main_bas,length,lat,lon,sex,ind.id) %>%
              rename(obs.sl = length))
  #left_join(var.lats) %>%
  #mutate(ecoreg = fct_reorder(as.factor(ecoreg), -m.lat))

dfp %>%
  #filter(is.na(sex)) %>%
  ggplot() +
  geom_abline(slope = 1, color = "darkgrey") +
  geom_point(aes(x = obs.sl, y = exp(pred.sl), color = sex), shape = 4, alpha = 0.7, show.legend = FALSE) +
  facet_wrap(~main_bas) +
  theme_minimal() +
  labs(y="predicted", x="observed")
#ggsave(paste0(home, "/report/silv_po.png"), width = 17, height = 14, units = "cm")


```

### logistic p ~ length

Solve for z = 0 (i.e. when p = 0.5), which means that we want to get the observed length in the expression for $z$; 

$0 = alpha[su] + b_{0,su} * length_sc + b_{1,su} * age_sc$

Conditioned on mean age (age_sc = 0), this becomes:

$L50_{su} = -alpha[su] / b_{0,su}$

```{r}
# Adding covariate effects to the L50
# First make mean observed lat and inland/marine to each su
df_sucov <- df.sweel3 %>%
  summarise(lat_scmean = mean(lat_sc), 
            .by = su) %>%
  left_join(df.sweel3 %>% distinct(su,spat.unit,marine,sb,balance)) %>%
mutate(coast = if_else(marine == 1, "coast","inland"),
         sb = if_else(sb == 1, "Westcoast","Baltic Sea"))

df_coef <- s.samples %>%
  as_draws() %>%
  spread_draws(alpha[su], logb0[su]) %>%
  select(-.iteration,-.chain) %>%
  left_join(df_sucov) %>%
  left_join(s.samples %>%
              as_draws() %>%
              spread_draws(b1,b2,b3,b4[sb]), by = c("sb",".draw"))

df_data <- df.sweel3 %>%
  select(length,ind.id,su,sb,marine,lat_sc) %>%
  left_join(s.samples %>%
              as_draws() %>%
              spread_draws(age_sc[ind.id],logweight[ind.id]) %>%
              median_qi()
              ) %>%
  cross_join(s.samples %>%
              as_draws() %>%
              spread_draws(ab[par]) %>%
              median_qi() %>%
              pivot_wider(values_from = c(.lower,.upper,ab), names_from = par)) %>%
  mutate(logw_mu = ab_1 + ab_2*log(length),
         le_cren = exp(logweight)/exp(logw_mu)) %>%
  select(ind.id,le_cren,age_sc,su,marine,lat_sc)
  
lengths <- seq(min(df.sweel3$length_sc), max(df.sweel3$length_sc), length.out = 100)
res <- vector("list", nsu)

for(j in 1:nsu){
  
  dat_j <- df_data %>% filter(su == j)
  coef_j <- df_coef %>% filter(su == j)
  
  res_j <- lapply(lengths, function(l){
    
    X_ind <- cbind(dat_j$age_sc, dat_j$le_cren, dat_j$marine)  # n_ind x 3
    #X_ind = as.matrix(dat_j %>% select((age_sc,le_cren, marine))  # n_ind x 3
    B_draws <- cbind(coef_j$b1, coef_j$b2, coef_j$b4)  # n_draws x 3
    z_unit <- coef_j$alpha + exp(coef_j$logb0)*l + coef_j$b3*dat_j$lat_sc[1]
  
    z_mat <- B_draws %*% t(X_ind) + as.vector(z_unit)  # n_draws x n_individuals
  
    p_mat      <- 1 / (1 + exp(-z_mat))
    p_marginal <- rowMeans(p_mat)  # average over individuals: n_draws vector
  
    data.frame(
      su        = j,
      length_sc = l,
      p_mean    = mean(p_marginal),
      p_lo      = quantile(p_marginal, 0.025),
      p_up      = quantile(p_marginal, 0.975)
      )
    })
  
  res[[j]] <- bind_rows(res_j)
}

# for smaller datasets, dfl is both draws and data combined
# for(j in 1:nsu) {
#   tmp = dfl %>%
#     filter(su == j) %>%
#     expand_grid(length_sc = seq(min(df.sweel3$length_sc, na.rm = TRUE), max(df.sweel3$length_sc, na.rm = TRUE), by = 0.1)) %>%
#     p = 1 / (1 + exp(-(alpha + exp(logb0)*length_sc + b1*age_sc + b2*le_cren + b3*lat_scmean + b4*marine))) %>%
#     group_by(length_sc) %>%
#     median_qi(p) %>%
#     mutate(sup = j) %>%
#     bind_rows(t)
#   rm(tmp, tmp2)
#   gc()
#   }
dfs <- s.samples %>%
  as_draws() %>%
  spread_draws(silver[ind.id]) %>%
  median_qi() %>%
  left_join(df.sweel3 %>% distinct(ind.id,su,length)) %>%
  left_join(df_sucov) %>%
  ungroup() %>%
  mutate(coast = if_else(marine == 1, "coast","inland"),
         sb = if_else(sb == 1, "Westcoast","Baltic Sea"))

# by each spatial unit
results %>%
  bind_rows(results) %>%
  mutate(length = length_sc*sd(df.sweel3$length, na.rm = TRUE) + mean(df.sweel3$length, na.rm = TRUE)) %>% 
  left_join(df_sucov %>% select(-sb)) %>%
  ggplot() +
  geom_line(aes(length, p_mean)) +
  geom_ribbon(aes(x = length, y = p_mean, ymin = p_lo, ymax = p_hi, alpha = 0.5, color = NA)) +
  #, group = spat.unit), fill = factor(balance)), alpha = 0.5, color = NA) + 
  geom_point(data = dfs %>% left_join(df_sucov %>% select(-sb)), 
             aes(length, y = silver, color = factor(balance)), 
             alpha = 0.3, size = 0.5) +
  facet_wrap(~su)

# by habitat and sea basin
results %>%
  bind_rows(results) %>%
  left_join(df_sucov) %>%
  mutate(length = length_sc*sd(df.sweel3$length, na.rm = TRUE) + mean(df.sweel3$length, na.rm = TRUE)) %>% 
  left_join(df_sucov %>% select(-sb)) %>%
  ggplot() +
  geom_line(aes(x = length, y = p_mean, group = lat_scmean, color = lat_scmean)) +
  geom_point(data = dfs, aes(length, y = silver, group = lat_scmean, color = lat_scmean), show.legend = FALSE, alpha = 0.3, shape = 124, size = 1.5) +
  facet_wrap(coast~sb) +
  scale_color_viridis_c()

# test 1, not correct
# dfzl <- s.samples %>%
#   as_draws() %>%
#   spread_draws(alpha[su], logb0[su]) %>%
#   median_qi() %>%
#   left_join(df_sucov) %>%
#   left_join(s.samples %>%
#               as_draws() %>%
#               spread_draws(b3,b4[sb]) %>%
#               median_qi(), by = "sb") %>%
#   expand_grid(length_sc = seq(min(df.sweel3$length_sc, na.rm = TRUE), max(df.sweel3$length_sc, na.rm = TRUE), by = 0.075)) %>%
#   #left_join(df.sweel3 %>% distinct(spat.unit,su, sb)) %>%
#   ungroup() %>%
#   # mutate(p_mean = 1 / (1 + exp(-(alpha + exp(logb0)*length_sc + b3*lat_scmean + b4*marine))),
#   #        p_upper = 1 / (1 + exp(-(alpha.upper + exp(logb0.upper)*length_sc + b3.upper*lat_scmean + b4.upper*marine))),
#   #        p_lower = 1 / (1 + exp(-(alpha.lower + exp(logb0.lower)*length_sc + b3.lower*lat_scmean + b4.lower*marine)))
#          mutate(p_mean  = 1 / (1 + exp(-(alpha       + exp(logb0) *length_sc))),
#                 p_upper = 1 / (1 + exp(-(alpha.upper + exp(logb0.upper)*length_sc))),
#                 p_lower = 1 / (1 + exp(-(alpha.lower + exp(logb0.lower)*length_sc)))
#          )

# test 2, using the posterior and propagating error by calc. L50 by draw
# dfzl <- s.samples %>%
#   as_draws() %>%
#   spread_draws(alpha[su], logb0[su]) %>%
#   select(-.iteration,-.chain) %>%
#   left_join(df_sucov) %>%
#   left_join(s.samples %>%
#               as_draws() %>%
#               spread_draws(b1,b2,b3,b4[sb]), by = c("sb",".draw")) %>%
#   expand_grid(length_sc = seq(min(df.sweel3$length_sc, na.rm = TRUE), max(df.sweel3$length_sc, na.rm = TRUE), by = 0.075)) %>%
#   mutate(p = 1 / (1 + exp(-(alpha + exp(logb0)*length_sc + b3*lat_scmean + b4*marine)))) %>%
#   group_by(su,length_sc) %>%
#   median_qi(p) %>%
#   left_join(df_sucov) %>%
#   mutate(coast = if_else(marine == 1, "coast","inland"),
#          sb = if_else(sb == 1, "Westcoast","Baltic Sea"))
#   
# dfzl %>%
#   #filter(balance == 1) %>%
#   mutate(length = length_sc*sd(df.sweel3$length, na.rm = TRUE) + mean(df.sweel3$length, na.rm = TRUE)) %>%
#   ggplot() +
#   geom_line(aes(x = length, y = p, group = spat.unit, color = factor(balance)), linewidth = 0.2) +
#   geom_ribbon(aes(x = length, y = p, ymin = .lower, ymax = .upper, group = spat.unit, fill = factor(balance)), 
#               alpha = 0.5, color = NA) +
#   geom_point(data = dfs, 
#              aes(length, y = silver, color = factor(balance)), 
#              alpha = 0.3, size = 0.5) +
#   facet_wrap(~spat.unit)
# 
# dfzl %>%
#   mutate(length = length_sc*sd(df.sweel3$length, na.rm = TRUE) + mean(df.sweel3$length, na.rm = TRUE)) %>%
#   ggplot() +
#   geom_line(aes(x = length, y = p, group = lat_scmean, color = lat_scmean)) +
#   geom_point(data = dfs, aes(length, y = silver, group = lat_scmean, color = lat_scmean), show.legend = FALSE, alpha = 0.3, shape = 124, size = 1.5) +
#   facet_wrap(coast~sb) +
#   scale_color_viridis_c()

```

### logistic p ~ age
Calculate for basin specific L50.

Solve for z = 0 (i.e. when p = 0.5), which means that we want to get the observed length in the expression for $z$; 

$0 = alpha[su] + b_{0,su} * length_sc + b_{1,su} * age_sc$

Conditioned on mean age (length = 0), this becomes:

$A50_{su} = -alpha[su] / b_{1,su}$

```{r}
s.samples %>%
  as_draws() %>%
  spread_draws(alpha[su], b1) %>%
  median_qi()

# using alpha, b0 which is per basin 
dfza <- s.samples %>%
  as_draws() %>%
  spread_draws(alpha[su], log_b0[su], b1) %>%
  median_qi() %>%
  expand_grid(age_sc = seq(min(df.sweel3$age_sc, na.rm = TRUE),max(df.sweel3$age_sc, na.rm = TRUE), by = 0.05)) %>%
  left_join(df.sweel3 %>% distinct(spat.unit,su)) %>%
  ungroup() %>%
  mutate(p_mean = 1 / (1 + exp(-(alpha + exp(log_b0)*0 + b1*age_sc))))

ssu <- sample(df.sweel3$su, 20, replace = FALSE)
dfza %>%
  filter(#sex == 0,
         su %in% ssu) %>%
  ggplot() +
  geom_line(aes(x = age_sc, y = p_mean, color = factor(su)), show.legend = FALSE) 
  #geom_point(data = df.sweel3 %>% filter(su %in% ssu), aes(age_sc, y = silver_dur, color = factor(spat.unit)),show.legend = FALSE, alpha = 0.3) 
  facet_wrap(~spat.unit) 

dfza %>%
  filter(sex == 0) %>%
  ggplot() +
  geom_line(aes(x = age_sc, y = p_mean, color = factor(su)), show.legend = FALSE)
```

```{r}
dfp <- silv.samples %>%
  spread_draws(s.mu[i]) %>%
  left_join(data.silv %>% 
              select(ecoreg,sex,fi_year) %>% 
              mutate(i = row_number()), 
            by = "i") %>%
  group_by(.draw, ecoreg, sex, fi_year) %>%
  summarise(pred.sl = median(s.mu), .groups = "drop") %>%
  left_join(var.lats) %>%
  mutate(ecoreg = fct_reorder(as.factor(ecoreg), m.lat),
         sex = if_else(sex == 0, "female", "male")) %>%
  mutate(m_pred = median(pred.sl), .by = c(sex))

n.labs <- data.silv %>%
  summarise(n = n(), .by = c(ecoreg, sex)) %>%
  mutate(sex = if_else(sex==1, "female", "male"))

dfp %>% summarise(mexp = exp(median(pred.sl)), .by= sex)
dfp %>% ungroup() %>% summarise(exp(median(pred.sl)), .by= sex)

dfp %>%
  ggplot() +
  geom_boxplot(aes(x = exp(pred.sl), y = ecoreg), outliers = FALSE, show.legend = FALSE) +
  geom_vline(aes(xintercept = exp(m_pred)), color = "darkgrey", size = 0.5) +
  facet_wrap(~sex, scales = "free_x") +
  theme_minimal() +
  labs(x = "Length at silvering [mm]", y = "ecoregion")
ggsave(paste0(home, "/report/silv_ERl.png"), width = 17, height = 14, units = "cm")

```

### Length over space

```{r, cache = FALSE}
# L50 by points
dfs <- s.samples %>%
  as_draws() %>%
  spread_draws(alpha[su], log_b0[su], sep = ",") %>%
  median_qi() %>%
  mutate(L50 = -alpha/exp(log_b0)) %>%
  # gather_draws(alpha[mb], bs[er], bt, sep = ",") %>%
  left_join(df.sweel3 %>% distinct(spat.unit,su,lat,lon))# %>%

# by catchment
dfsca <- ca_sf %>%
  mutate(spat.unit = as.character(main_bas)) %>%
  #mutate(main_bas = as.factor(main_bas)) %>%
  left_join(dfs %>% distinct(su,L50,spat.unit)) %>%
  filter(!is.na(mb)) #%>%
  mutate(pred.length = exp(.value))
  
plot_map_Nord +
  geom_sf(data = dfs, aes(fill = L50), color = NA,  alpha = 0.5) +
  scale_fill_viridis_c() +
  geom_point(data = df.sweel2 %>% drop_na(lat,lon) %>% 
               add_utm_columns(ll_names = c("lon", "lat"), utm_crs = 32633), aes(X*1000, Y*1000), #color = , 
             size = 0.1) 
  
# # ungroup() %>%
#   # mutate(female = alpha + bt*mt,
#   #        male = alpha + bs + bt*mt,
#   #        year = yr + 1997) %>%
#   pivot_longer(cols = c("female","male"), names_to = "sex", values_to = "length") %>%
#   mutate(length = exp(length)) %>%
#   mutate(m_length = median(length), .by = c(year,sex)) %>%
#   left_join(var.lats) %>%
#   mutate(ecoreg = fct_reorder(ecoreg, -m.lat)) 
# 
# dfp %>%
#   filter(sex == "female") %>%
#   ggplot() +
#   geom_boxplot(aes(y = length, x = year, group = year), color = "#F8766D", outliers = FALSE) +
#   geom_line(aes(y = m_length, x = year), color = "#F8766D", alpha = 0.4) +
#   facet_wrap(~ecoreg, scales = "free_y") +
#   theme_minimal() +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1, size = rel(.75))) +
#   labs(x = "year")
# ggsave(paste0(home, "/report/silv_lf.png"), width = 17, height = 14, units = "cm")
# 
# dfp %>%
#   filter(sex == "male") %>%
#   ggplot() +
#   geom_boxplot(aes(y = length, x = year, group = year), color = "#00BFC4", outliers = FALSE) +
#   geom_line(aes(y = m_length, x = year), color = "#00BFC4", alpha = 0.4) +
#   facet_wrap(~ecoreg, scales = "free_y") +
#   theme_minimal() +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1, size = rel(.75))) +
#   labs(x = "year")
# ggsave(paste0(home, "/report/silv_lm.png"), width = 17, height = 14, units = "cm")  

```

### Length distribution per su

```{r}
pis <- s.samples %>%
  as_draws() %>%
  #spread_draws(p[ind.id], length_sc[ind.id], age_sc[ind.id], weight_sc[ind.id], silver[ind.id]) %>%
  spread_draws(p[ind.id], age_sc[ind.id], logweight[ind.id], silver[ind.id]) %>%
  left_join(df.sweel3 %>% select(su,spat.unit,ind.id,sb,balance,marine,sea_basin,length), by = "ind.id") 

# Weighting has very little effect on the density distribution:
pis %>%
  #mutate(length = length_sc * sd(df.sweel3$length, na.rm = TRUE) + mean(df.sweel3$length, na.rm = TRUE)) %>%
  filter(silver == 1) %>%
  ggplot() + 
  geom_density_ridges(aes(x = length, y = sb, weight = p, fill = factor(sb)), color = NA, alpha = 0.5) +
  geom_density_ridges(aes(x = length, y = sb, color = factor(sb)), fill = NA) +
  labs(caption = "filled is weigthed by p")

# density by balanced sunits and sb 
p3 <- pis %>%
  mutate(length = length_sc * sd(df.sweel3$length, na.rm = TRUE) + mean(df.sweel3$length, na.rm = TRUE),
         sb = if_else(sb == 1, "Westcoast", "Baltic Sea")
         ) %>%
  filter(silver == 1,
         balance == 1
         ) %>%
  ggplot(aes(x = length, y = spat.unit, weight = p, fill = factor(sb))) + 
  geom_density_ridges(alpha = 0.8, color = NA)

pis %>%
  left_join(df.sweel3 %>% select(ind.id,marine), by = "ind.id") %>%
  mutate(habitat = if_else(marine == 1, "coast","inland"),
         sb = if_else(sb == 1, "Westcoast","Baltic Sea"),
         #length = length_sc * sd(df.sweel3$length, na.rm = TRUE) + mean(df.sweel3$length, na.rm = TRUE),
         ) %>%
  filter(silver == 1) %>%
  ggplot(aes(x = length, y = sb, weight = p, fill = habitat), color = NA) + 
  geom_density_ridges(alpha = 0.5)

p3 / p4 + plot_layout(ncol = 1, guides = "collect")

plot_map_Nord +
  geom_point(data = df.sweel3 %>% 
               filter(su %in% c(6,27)) %>%
               filter(balance == 1) %>% add_utm_columns(ll_names = c("lon", "lat"), utm_crs = 32633), aes(X*1000, Y*1000, color = sea_basin)) +
  geom_sf(data = ca_sf2, aes(fill = factor(main_catch)),  alpha = 0.3)

# 6, 27,59
```

### Posterior predictive checks
```{r}
# visual classifier
# compare means by spatial unit
tmps <- as_draws(s.samples) %>%
  #subset_draws(variable = c("visual_rep")) %>%
  spread_draws(visual_rep[ind.id]) %>%
  median_qi() %>%
  left_join(df.sweel3 %>% select(su,spat.unit,ind.id), by = "ind.id") 

tmps %>%
  summarise(mean_est_vis = mean(visual_rep, na.rm = TRUE), .by = spat.unit) %>%
  left_join(df.sweel3 %>% summarise(obs_vis = mean(silver_vis, na.rm = TRUE), .by = spat.unit), by = "spat.unit")

# plot densities of 0/1 
tmps %>%
  ggplot() +
  geom_density(aes(visual_rep, group = ), color = "red") +
  geom_density(data = df.sweel3, aes(silver_vis)) +
  labs(title = "red is model")

# Durif classifier
tmpd <- as_draws(s.samples) %>%
  #subset_draws(variable = c("visual_rep")) %>%
  spread_draws(measured_rep[ind.id]) %>%
  median_qi() %>%
  left_join(df.sweel3 %>% select(su,spat.unit,ind.id), by = "ind.id") 

tmpd %>%
  summarise(mean_est_dur = mean(measured_rep, na.rm = TRUE), .by = spat.unit) %>%
  left_join(df.sweel3 %>% summarise(obs_dur = mean(silver_dur, na.rm = TRUE), .by = spat.unit), by = "spat.unit")

tmpd %>%
  ggplot() +
  geom_density(aes(measured_rep), color = "red") +
  geom_density(data = df.sweel3, aes(silver_vis)) +
  labs(title = "red is model")
  ggplot() +
  geom_density(aes(measured_rep)) +
  geom_density(data = df.sweel3, aes(silver_dur))

# Test
# Disagreement rate between classifiers — the model's ability to reproduce the observed rate of V≠M is a direct check on Se/Sp estimation
#Prevalence by population — check that simulated stage prevalence matches observed classifier prevalence within each population mb
#Prevalence along the length gradient — bin individuals by length and compare observed vs simulated classifier rates, which checks the logistic curve f
```

### The NA stage III:s

```{r}
tmp3s <- s.samples %>%
  as_draws() %>%
  spread_draws(silver[ind.id]) %>%
  median_qi() %>%
  left_join(df.sweel3 %>% select(visual_stage,durif_stage,ind.id,su), by = "ind.id") 

tmp %>%
  filter(durif_stage %in% c("FII","FIII","FIV","FV")) %>%
  summarise(n = n(), .by = c(silver, durif_stage, visual_stage)) %>%
  arrange(-silver, -n) %>%
  rename(estimated = silver)

```

# test distributions
```{r}
  mu_a_gl ~ dnorm(msa, sda)
  sd_a_gl ~ dgamma(shape = 1.5, rate = 1.5 / sda)
  #mu_b_gl ~ dnorm(log(mb0), sd = log(sdb0))
  mu_b_gl ~ dnorm(log(mb0), sd = 0.5)
  #sd_b_gl ~ dgamma(shape = 1.5, rate = 1.5 / sdb0) # to weak 
  sd_b_gl ~ dgamma(shape = 2, rate = 4 / sdb0) 
  
  sd_a_su ~ dgamma(shape = 2, rate = 2cx)   # sd for su-level
  
  
ggplot() +
  #stat_function(fun = dgamma, 
  #            args = list(shape = 1000, rate = 2000)) +
  #stat_function(fun = dgamma, 
  #            args = list(shape = 2, rate = 12), color = "blue") +
  stat_function(fun = dexp, 
              args = list(7), color = "blue") +
  stat_function(fun = dexp, 
              args = list(3)) +
  xlim(0,5)
```

```{r}
tmpab <- 
  subset_draws(as_draws(s.samples), variable = c("ab")) %>%
  summarise_draws(c("mean","sd")) %>%
  mutate(variable = case_when(variable == "ab[1]" ~ "a",
                              variable == "ab[2]" ~ "b")) %>%
  pivot_wider(names_from = variable, values_from = c(mean,sd))
                           
df.sweel3 %>%
  select(length,weight,su) %>%
  drop_na(length,weight) %>%
  bind_cols(tmpab) %>%
  mutate(logweight_mu = mean_a + mean_b*log(length),
         le_cren = weight / exp(logweight_mu)) %>%
  filter(le_cren < 0.001) %>%
  ggplot() +
  #geom_point(aes(length, weight))
  geom_density(aes(le_cren
                  ))
```

