---
title: "Swedish eel silvering model summary"
subtitle: "SWEELHT"
author: "Viktor Thunell" 
date: "`r format(Sys.time(), '%d %B, %Y')`"
format: 
  html:
    code-fold: true
    code-summary: "Show code"
    page-layout: full
    embed-resources: true
    toc: true
knitr: 
  opts_chunk:
    fig.align: center
    out-width: 80%
editor: source
execute: 
  echo: true
  eval: true
  cache: true
---

# Latent classs logistic regression for eel migration probability

## Load libraries

```{r}
#| message: false
#| warning: false
#| cache: false

# Load libraries, install if needed
pkgs <- c("tidyverse", "tidylog", "devtools","viridis","sf","patchwork","posterior","ggridges","arrow", "tidybayes","bayesplot","nimble","here","reshape2","sdmTMB","broom")

if(length(setdiff(pkgs,rownames(installed.packages()))) > 0){
    install.packages(setdiff(pkgs, rownames(installed.packages())), dependencies = T)
  }

invisible(lapply(pkgs, library, character.only = T))

devtools::source_url("https://raw.githubusercontent.com/VThunell/diasp-lht/main/R/functions/map-plot.R")
options(ggplot2.continuous.colour = "viridis")
theme_set(theme_minimal()) 

# Set path
home <- here::here()
```

## Read data

```{r}
#| message: false
#| warning: false
#| cache: false

## Read data
#sweels <- readRDS(file = paste0(home, "/data/sweel_data_2026-06-24.RData"))
## Read data for 4f
sweels <- readRDS(file = paste0(home, "/data/sweel_data_2026-09-10.RData"))
```

```{r}
# lets keep the blue ones as females:
sweels %>%
  mutate(ld = if_else(length > 650 & is.na(sex),"pf","na")) %>%
  ggplot() +
  geom_point(aes(length, weight, color = ld)) +
  facet_wrap(~sex)
```

## Silvering logistic regression model

### Define data

```{r}
# filter females and keep only weights and lengths
df.sweel <- sweels %>%
  filter(!is.na(length)) %>%
  mutate(weight = if_else(weight == 0, NA, weight)) %>%
  # remove rows where used variables have at least two non NAs
  filter(rowSums(!is.na(select(., visual_stage, weight, length, age))) >= 2) %>%
  #keep the more than 50 cm long NA sex as females
  mutate(sex = if_else(is.na(sex) & length > 650, "female", sex)) %>%
  filter(sex == "female")

# make the con,eye and pec index 
abw <- df.sweel %>%
  drop_na(weight,length) %>%
  lm(data = ., log(weight) ~ log(length)) %>%
  tidy()

abe <- df.sweel %>%
  drop_na(diameter_eye,length) %>%
  lm(data = ., log(diameter_eye) ~ log(length)) %>%
  tidy()

abp <- df.sweel %>%
  drop_na(length_pectoral,length) %>%
  lm(data = ., log(length_pectoral) ~ log(length)) %>%
  tidy()

df.sweel2 <- df.sweel %>% 
  mutate(silver_vis = case_when(visual_stage == "Blankål" ~ 1,
                                visual_stage == "Halvblank ål" ~ NA,
                                visual_stage %in% c("Gulål","Yngel","Juvenil") ~ 0,
                                .default = NA),
         length_sc = (length - mean(length, na.rm = TRUE)) / sd(length, na.rm = TRUE),
         age_sc = (age - mean(age, na.rm = TRUE)) / sd(age, na.rm = TRUE),
         con_index = weight/exp(abw$estimate[1] + abw$estimate[2]*log(length)),
         pec_index = length_pectoral/exp(abp$estimate[1] + abp$estimate[2]*log(length)),
         eye_index = diameter_eye/exp(abe$estimate[1] + abe$estimate[2]*log(length))) %>%
  # some possible glasss eels are distorting the con index, remove these
  filter(con_index < 100 | is.na(con_index)) %>%
  mutate(con_index_sc = (con_index - mean(con_index, na.rm = TRUE)) / sd(con_index, na.rm = TRUE),
         pec_index_sc = (pec_index - mean(pec_index, na.rm = TRUE)) / sd(pec_index, na.rm = TRUE),
         eye_index_sc = (eye_index - mean(eye_index, na.rm = TRUE)) / sd(eye_index, na.rm = TRUE),
         sb = case_when(sea_basin == "VH" ~ 1,
                        !sea_basin == "VH" ~ 2),
         #sb = as.integer(factor(sea_basin)),
         sa = as.integer(factor(sea_area)),
         spat_unit = case_when(marine == "yes" ~ sea_basin,
                               !is.na(haro) & marine == "no" ~ haro,
                               is.na(haro) & marine == "no" ~ as.character(main_catch),
                               .default = NA),
         #spat_unit = if_else(marine == "yes", sb, main_catch),
         su = as.integer(factor(spat_unit)),
         habitat = if_else(marine == "yes", 1,0),
         lat_sc = (lat - mean(lat, na.rm = TRUE)) / sd(lat, na.rm = TRUE),
         ) %>%
  # still one individual with very unlikely length weight combo, remove
  #filter(!(con_index > 4 & !is.na(con_index))) %>%
  mutate(silver_vis_orig = silver_vis,
         ind_id = row_number())

df.sweel3 <- df.sweel2 %>%
  mutate(silvratio = mean(silver_vis == 1, na.rm = TRUE), .by = spat_unit) %>%
  mutate(balance = if_else(silvratio < 0.1 | silvratio > 0.9 | is.na(silvratio), 0, 1)) 

```

### Plot silvering data
```{r}
#In space
plot_map_Nord +
  #geom_sf(data = ca_sf %>% filter(main_catch %in% df_cobaca_inland$main_catch), aes(fill = factor(main_catch)), color = NA,  alpha = 0.3)  +
  geom_point(data = df.sweel3 %>%
               mutate(Habitat = if_else(marine == "no", "Inland", "Marine")) %>%
               add_utm_columns(ll_names = c("lon", "lat"), utm_crs = 32633), aes(X*1000, Y*1000, color = Habitat), size = 0.1) +
  guides(color = guide_legend(title.position = "top")) +
  labs(color = "") +
  theme(legend.position = "inside", legend.position.inside = c(0.18, 0.80))
ggsave(paste0(home, "/figures/map_data.pdf"), width = 18, height = 13, units = "cm")

# density of length by visual classifier
df.sweel3 %>%
  filter(balance == 1) %>%
  ggplot() +
  geom_density(aes(length, fill = factor(silver_vis)), color = NA, alpha = 0.3) +
  facet_wrap(~spat_unit, scales = "free")
  
df.sweel3 %>%
  filter(balance == 0) %>%
  ggplot() +
  geom_density(aes(length, fill = factor(silver_vis)), color = NA, alpha = 0.3) +
  facet_wrap(~spat_unit)

# correlation between covariates
df.sweel3 %>%
  select(length_sc, age_sc, con_index_sc, pec_index_sc, eye_index_sc) %>%
  drop_na(length_sc, age_sc, con_index_sc, pec_index_sc, eye_index_sc) %>%
  ggpairs(
  lower = list(continuous = wrap("smooth", alpha = 0.3, color = "blue")), # Scatter plots with trendline
  diag = list(continuous = wrap("densityDiag", fill = "gray")),           # Density plots on diagonal
  upper = list(continuous = wrap("cor", size = 4), progress = FALSE)
) 

# relationship between length, indices and visual classifer
df.sweel3 %>%
  ggplot(aes(length, silver_vis)) +
  geom_point(size = 0.1, alpha = 0.3) +
  facet_wrap(~su)

a<-df.sweel3 %>%
  ggplot() +
  geom_density(aes(con_index_sc, fill = factor(silver_vis)), alpha = 0.5, color = NA, show.legend = FALSE) 
b<-df.sweel3 %>%
  ggplot() +
  geom_density(aes(eye_index_sc, fill = factor(silver_vis)), alpha = 0.5, color = NA, show.legend = FALSE) 
c<-df.sweel3 %>%
  ggplot() +
  geom_density(aes(pec_index_sc, fill = factor(silver_vis)), alpha = 0.5, color = NA) 

a/b/c

# relationship between length and le_cren, latitude
df.sweel3 %>%
  ggplot(aes(con_index_sc, length, color = factor(silver_vis))) +
  geom_point(size = 0.1, alpha = 0.3, show.legend = FALSE) +
  
df.sweel3 %>%
  ggplot(aes(eye_index_sc, length, color = factor(silver_vis))) +
  geom_point(size = 0.1, alpha = 0.3, show.legend = FALSE) +
  
df.sweel3 %>%
  ggplot(aes(pec_index_sc, length, color = factor(silver_vis))) +
  geom_point(size = 0.1, alpha = 0.3, show.legend = FALSE) +
  
df.sweel3 %>%
  ggplot(aes(lat_sc, length, color = factor(silver_vis))) +
  geom_point(size = 0.1, alpha = 0.3, show.legend = FALSE) +
  geom_smooth(method = "lm", se = TRUE)

# relationship between length and silver
df.sweel3 %>%
  ggplot() +
  geom_histogram(aes(length, fill = factor(silver_vis)), alpha = 0.5)

# LHTs of Durif staged eels
df.sweel3 %>%
  filter(!is.na(durif_stage)) %>%
  ggplot(aes(length, factor(su), color = durif_stage)) +
  geom_boxplot(outlier.color = "grey",
               outlier.size = 0.5) +
df.sweel3 %>%
  filter(!is.na(visual_stage)) %>%
  ggplot(aes(length, factor(su), color = visual_stage)) +
  geom_boxplot(outlier.color = "grey",
               outlier.size = 0.5) 

df.sweel3 %>%
  filter(!is.na(durif_stage)) %>%
  ggplot(aes(age, factor(su), color = durif_stage)) +
  geom_boxplot(outlier.color = "grey",
               outlier.size = 0.5) 
df.sweel3 %>%
  filter(!is.na(durif_stage)) %>%
  ggplot(aes(weight, factor(su), color = durif_stage)) +
  geom_boxplot(outlier.color = "grey",
               outlier.size = 0.5) 
```
### n over space and time 

```{r}
df.sweel3 %>%
  summarise(count = n(), .by = c(su))

df.sweel3 %>%
  summarise(count = n(), .by = c(year, su)) %>%
  ggplot(aes(year, factor(su), fill = count)) +
  geom_tile(color = "gray30") +
  #geom_vline(xintercept = 1998, color = "darkgrey") +
  scale_fill_viridis_c() +
  labs(y = "", x = "year") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Source models / Load samples
```{r, cache = FALSE}
# Load samples
#sweelsilv.samples <- readRDS(paste0(home,"/data/samples/sweelsilv.samples_g2026-08-15.RData"))
sweelsilv.samples <- readRDS(paste0(home,"/data/samples/sweelsilv.samples_2026-09-09.RData"))
sweelsilv.samples$WAIC
# for simple 33029.01 - 20260507
# for spatial alpha 13956.79 - 20260507
# for spatial and hierachical alpha 13875.92 - 20260508
# for spatial b0 and hierachical spatial alpha 13570.31 - 20260508
# for spatial and hierachical alpha and b0 13506.59 - 20260511
# for spatial and hierachical alpha and b0 21417.19 - 20260511 (this adds data so incomparable to previous models)
# for spatial and hierachical alpha and b0 with two classifiers and se/sp: 8411.084
# 225708 # new data and model after meeting 4/6 (model 4, no condition)
#  234143.4 20260615 and then 231746.8 for the 4c model with 
s.samples <- sweelsilv.samples$samples

# # Or source model code and mcmc (load libs and data)
# t <- Sys.time()
# source(file = paste0(home,"/models_eel/model_eel_silvering.R"))
# Sys.time() - t # approx 4.5 hours with 10000 hmc samples
# WAIC based on combined log prob of silver_vis and _dur; d: 25569.66 (22946.08), e:26438.2
#401627.6 vs  395643 
# 395708 5/9
# 395827 9/9
# 334249.5 10/9

# unstable pWAIC values are all data nodes (silver_vis)
# pW <- bind_cols(pWAIC=sweelsilv.mcmcc$getWAICdetails(returnElements = TRUE)$pWAIC_elements,
#                 name=sweelsilv.model$getNodeNames(dataOnly = TRUE)) %>%
#   filter(pWAIC > 0.4)
# sweelsilv.model$getVarNames(nodes = pW %>% pull(name))

```

### Traces, ess, and ‘potential scale reduction factor’

Variables "s.sig","Ustar","U","mu.a.gl","sd.a.gl","alpha","bs","bt")
```{r, cache = FALSE}
variable.names(as_draws(s.samples)) %>%
  str_remove("\\[.*\\]") %>% unique()

node.sub <- subset_draws(as_draws(s.samples), variable = 
                           c("mu_a_gl","mu_b_gl"))
                           #c("alpha_sb","logb0_sb","mu_a_gl","mu_b_gl"))
node.sub[] %>% mcmc_trace() 
node.sub %>% summarise_draws(c("mean","median","sd","rhat","ess_bulk","ess_tail"))
#log odds intercept to prob 1/(1+exp(z)). 
node.sub %>% summarise_draws(c("mean"))
1/(1+exp(-15)) # a silver of mean length in the data (631 mm) has a 99,99 % prob to be a silver

node.sub <- subset_draws(as_draws(s.samples), variable = c("alpha"))
#node.sub[,,sample(1:dim(node.sub)[3], 16)] %>% mcmc_trace() # two chain
node.sub[,sample(1:dim(node.sub)[2], 16)] %>% mcmc_trace() # one chain
node.sub %>% summarise_draws(c("mean","median","sd","rhat","ess_bulk","ess_tail")) %>% arrange(-rhat)

node.sub <- subset_draws(as_draws(s.samples), variable = c("logb0"))
node.sub[,sample(1:dim(node.sub)[2], 16)] %>% mcmc_trace() # one chain
node.sub %>% summarise_draws(c("mean","median","sd","rhat","ess_bulk","ess_tail")) %>% arrange(-rhat) %>% mutate(mean = exp(mean))

node.sub <- subset_draws(as_draws(s.samples), variable = #c("sp_v","sp_d","se_v","se_d","b1","b2","b3","b4","ab","w_nu","w_sd"))
c("sp_v","se_v","bA","bC","bLa","bH","bP","bE"))
node.sub[] %>% mcmc_trace() # one chain
node.sub %>% summarise_draws(c("mean","median","sd","rhat","ess_bulk","ess_tail"))
  
node.sub <- subset_draws(as_draws(s.samples), variable = "p")
#node.sub[,,sample(1:dim(node.sub)[3], 16)] %>% mcmc_trace()
node.sub[,sample(1:dim(node.sub)[2], 36)] %>% mcmc_trace()
#node.sub %>% mcmc_trace() # one chain
node.sub[,sample(1:dim(node.sub)[2], 600)] %>% summarise_draws(c("median","sd","rhat","ess_bulk","ess_tail")) %>% arrange(-rhat)

node.sub <- subset_draws(as_draws(s.samples), variable = "silver_rep")
node.sub <- subset_draws(as_draws(s.samples), variable = "measured_rep")
node.sub[,sample(1:dim(node.sub)[2], 36)] %>% mcmc_trace()
#node.sub %>% mcmc_trace() # one chain
node.sub[,sample(1:dim(node.sub)[2], 600)] %>% summarise_draws(c("median","sd","rhat","ess_bulk","ess_tail")) %>% arrange(-rhat)

```

### Silvering variables

```{r}
# alpha_sb/beta0_sb (sea basin unit), baltic vs west coast
# s.samples %>%
#   as_draws() %>%
#   spread_draws(alpha[su]) %>%
#   ggplot() +
#   geom_density(aes(x = alpha, color = factor(su))) +
# 
# s.samples %>%
#   as_draws() %>%
#   spread_draws(logb0[sb]) %>%
#   ggplot() +
#   geom_density(aes(x = logb0_sb, color = factor(sb))) 

# alpha an b0 by su
s.samples %>%
  as_draws() %>%
  spread_draws(alpha[su]) %>%
  left_join(df.sweel3 %>% distinct(su,sb,balance)) %>%
  ungroup() %>%
  ggplot() +
  geom_boxplot(aes(x = alpha, y = su, color = factor(balance), group = factor(su)), outliers = FALSE) 

s.samples %>%
  as_draws() %>%
  spread_draws(logb0[su]) %>%
  left_join(df.sweel3 %>% distinct(su,sb,balance)) %>%
  ggplot() +
  geom_boxplot(aes(x = logb0, y = su, color = factor(balance), group = factor(su)), outliers = FALSE)

# correlation between alpha and log_b0 ?
s.samples %>%
  as_draws() %>%
  spread_draws(alpha[su], logb0[su]) %>%
  median_qi() %>%
  ggplot(aes(y = alpha, x = logb0, xmin = logb0.lower, xmax = logb0.upper, ymin = alpha.lower, ymax = alpha.upper)) +
  geom_pointinterval(orientation = "vertical", point_size = 0.8, linewidth = 0.5) +
  geom_pointinterval(orientation = "horizontal", point_size = 0.8, linewidth = 0.5) 

# conditional L50 for mean age_sc (b2,b3 and b4 equals zero)
# s.samples %>%
#   as_draws() %>%
#   spread_draws(alpha[su], logb0[su]) %>%
#   mutate(L50_sc = -alpha/exp(logb0),
#          L50 = L50_sc * sd(df.sweel3$length) + mean(df.sweel3$length)) %>%
#   left_join(df.sweel3 %>% distinct(su,sb,marine,balance)) %>%
#   filter(balance == 1) %>%
#   ggplot() +
#   geom_density(aes(L50, color = factor(su)), show.legend = FALSE) +
#   theme_minimal() +
#   xlim(-1000,1000)
#   
# # conditional A50 for mean age_sc (b1,b3 and b4 equals zero)
# s.samples %>%
#   as_draws() %>%
#   spread_draws(alpha[su], b1) %>%
#   mutate(A50_sc = -alpha/b1,
#          A50 = A50_sc * sd(df.sweel3$age) + mean(df.sweel3$age),
#          ) %>%
#   left_join(df.sweel3 %>% distinct(su,sb,marine,balanced)) %>%
#   filter(balance == 1) %>%
#   ggplot() +
#   geom_density(aes(A50_sc, color = factor(su)), show.legend = FALSE) +
#   theme_minimal() +
#   #xlim(0,1500)
#   xlim(-15,15) 
# 
# # 50% transition surface
# s.samples %>%
#   as_draws() %>%
#   spread_draws(alpha[su], log_b0[su], b1) %>%
#   median_qi() %>%
#   expand_grid(age_sc = seq(min(df.sweel3$age_sc, na.rm = TRUE), max(df.sweel3$age_sc, na.rm = TRUE), length.out = 10)) %>%
#   mutate(length_sc = (-alpha - b1 * age_sc)/exp(log_b0)) %>%
#   ggplot() +
#   geom_line(aes(age_sc, length_sc, color = factor(su))) 
  
# sensitivity and specificity 
# s.samples %>%
#   as_draws() %>%
#   #gather_draws(se_d,se_v,sp_d,sp_v) %>%
#   gather_draws(se_v,sp_v) %>%
#   ggplot() +
#   geom_density(aes(x = .value, color = .variable)) +
#   # stat_function(fun = dbeta, args = list(shape1 = 5, shape2 = 1), color = "black",linetype = "dashed") +
#   # stat_function(fun = dbeta, args = list(shape1 = 10, shape2 = 1), color = "grey",linetype = "dashed") 
#   xlim(0.9,1) 

# covariates 
s.samples %>%
  as_draws() %>%
  gather_draws(bA,bC,bE,bLa,bP,bH[sb],se_v,sp_v) %>%
  #gather_draws(bA,bC,bE,bLa,bP,se_v,sp_v) %>%
  mutate(Parameter = case_when(.variable == "bLa" ~ "Latitude",
                               .variable == "bA" ~ "Age",
                               .variable == "bC" ~ "Cond. index",
                               .variable == "bE" ~ "Eye index",
                               .variable == "bP" ~ "Pec. fin index",
                               .variable == "bH" ~ "Habitat",
                               .variable == "se_v" ~ "Sensitivity",
                               .variable == "sp_v" ~ "Specificity",
                               .default = NA) ) %>%
  rename(Estimate = .value) %>%
  ggplot() +
  geom_boxplot(aes(x = Estimate, y = Parameter, color = factor(sb)), outliers = FALSE)
  #geom_boxplot(aes(x = Estimate, y = Parameter),  outliers = FALSE)
ggsave(paste0(home, "/figures/param_estimate.pdf"), width = 18, height = 13, units = "cm")

# tmpcovs <- s.samples %>%
#   as_draws() %>%
#   gather_draws(p[ind_id],p_vis[ind_id],age_sc[ind_id],con_index_sc[ind_id],pec_index_sc[ind_id],eye_index_sc[ind_id]) %>%
#   median_qi() %>%
#   left_join(s.samples %>%
#               as_draws() %>%
#               spread_draws(p[ind_id]) %>% 
#               median_qi(), by = "ind_id") 
# 
# tmpcovs %>%
#   filter(!.variable %in% c("p","p_vis")) %>%
#   left_join(tmpcovs %>%
#               filter(!.variable %in% c("p","p_vis"))) %>%
#   ggplot() +
#   geom_point(aes(p, .value, color = .variable), alpha = 0.1) +
#   facet_wrap(~.variable, scales = "free")
```

### Gather draws by individual 

```{r}
df_inddra <- s.samples %>%
  as_draws() %>%
  spread_draws(p[ind_id],p_vis[ind_id],silver[ind_id],silver_vis[ind_id],con_index_sc[ind_id],age_sc[ind_id], pec_index_sc[ind_id],eye_index_sc[ind_id])


df_inddra_mq <- df_inddra %>%
  median_qi() %>%
  left_join(df.sweel3 %>% 
              distinct(ind_id,spat_unit,su,marine,sb,balance,sea_basin,length_sc))
```

### logistic p ~ length

Solve for z = 0 (i.e. when p = 0.5), which means that we want to get the observed length in the expression for $z$; 

$0 = alpha[su] + b_{L,su} * length_sc + b_{A,su} * age_sc + b_{H,su} * habitat + b_{1,su} * age_sc + b_{1,su} * age_sc$

Conditioned on mean age (age_sc = 0), this becomes:

$L50_{su} = -alpha[su] / b_{0,su}$

```{r}
# Adding covariate effects to the L50
# First make mean observed lat and inland/marine to each su
df_suinfo <- df.sweel3 %>%
  summarise(lat_scmean = mean(lat_sc), 
            .by = su) %>%
  left_join(df.sweel3 %>% distinct(su,habitat,sb,balance,spat_unit,sea_basin)) %>%
              mutate(coast = if_else(habitat == 1, "coast","inland"))

df_coef <- s.samples %>%
  as_draws() %>%
  spread_draws(alpha[su], logb0[su]) %>%
  select(-.iteration,-.chain) %>%
  left_join(df_suinfo %>% select(su,sb)) %>%
  left_join(s.samples %>%
              as_draws() %>%
              #spread_draws(bA,bC,bH,bE,bP,bLa), by = ".draw")
              spread_draws(bA,bC,bH[sb],bE,bP,bLa), by = c("sb",".draw"))

df_data <- df.sweel3 %>%
  select(length_sc,ind_id,su,sb,habitat,lat_sc) %>%
  left_join(df_inddra_mq) %>%
              #select(age_sc,con_index_sc,pec_index_sc,eye_index_sc,ind_id)) %>%
  # left_join(s.samples %>%
  #             as_draws() %>%
  #             spread_draws(age_sc[ind_id],con_index_sc[ind_id],pec_index_sc[ind_id],eye_index_sc[ind_id]) %>%
  #             median_qi()
  #             ) %>%
  select(length_sc,ind_id,con_index_sc,age_sc,su,habitat,lat_sc,pec_index_sc,eye_index_sc)
```

#### length logistic curves
```{r}
nsu <- length(unique(df.sweel3$su))
lengths <- c(seq(min(df.sweel3$length_sc), max(df.sweel3$length_sc), length.out = 100),0)
res_l <- vector("list", nsu)

for(j in 1:nsu){
  
  dat_j <- df_data %>% filter(su == j)
  coef_j <- df_coef %>% filter(su == j)
  
  res_j <- lapply(lengths, function(l){
    
    X_ind <- cbind(dat_j$age_sc, dat_j$con_index_sc, dat_j$eye_index_sc, dat_j$pec_index_sc, dat_j$habitat, dat_j$lat_sc)
    B_draws <- cbind(coef_j$bA, coef_j$bC, coef_j$bE, coef_j$bP, coef_j$bH, coef_j$bLa)
    z_unit <- coef_j$alpha + exp(coef_j$logb0)*l
  
    z_mat <- B_draws %*% t(X_ind) + as.vector(z_unit)  # n_draws x n_individuals
    
    p_mat      <- 1 / (1 + exp(-z_mat))
    p_marginal <- rowMeans(p_mat) # average over individuals: n_draws vector
  
    data.frame(
      su        = j,
      length_sc = l,
      p_mean    = mean(p_marginal),
      p_lo      = quantile(p_marginal, 0.025),
      p_up      = quantile(p_marginal, 0.975)
      )
    })
  
  res_l[[j]] <- bind_rows(res_j)
}

# by habitat and sea basin
bind_rows(res_l) %>%
  mutate(length = length_sc*sd(df.sweel3$length, na.rm = TRUE) +
           mean(df.sweel3$length, na.rm = TRUE)) %>% 
  left_join(df_suinfo) %>%
  ggplot() +
  geom_line(aes(x = length, y = p_mean, group = lat_scmean, color = lat_scmean, alpha = balance)) +
  geom_point(data = df_inddra_mq %>% 
               left_join(df.sweel3 %>% distinct(ind_id,silver_vis)) %>%
               left_join(df_suinfo %>% select(-sb)) %>%
                 mutate(length = length_sc*sd(df.sweel3$length, na.rm = TRUE) +
           mean(df.sweel3$length, na.rm = TRUE)),
             aes(length, y = silver_vis, group = lat_scmean, color = lat_scmean, alpha = balance), 
             show.legend = FALSE, alpha = 0.3, shape = 124, size = 1.5) +
  facet_wrap(~sea_basin) +
  scale_color_viridis_c()
ggsave(paste0(home, "/figures/L50_hab_basin.pdf"), width = 18, height = 13, units = "cm")

# by each spatial unit
bind_rows(res_l) %>%
  mutate(length = length_sc*sd(df.sweel3$length, na.rm = TRUE) + mean(df.sweel3$length, na.rm = TRUE)) %>% 
  left_join(df_suinfo %>% select(-sb)) %>%
  ggplot() +
  geom_line(aes(length_sc, p_mean, color = sea_basin)) +
  geom_ribbon(aes(x = length_sc, y = p_mean, ymin = p_lo, ymax = p_up), alpha = 0.3, color = NA) +
  geom_point(data = df_inddra_mq %>% left_join(df.sweel3 %>% distinct(ind_id,silver_vis)) %>% left_join(df_suinfo %>% select(-sb)),
  #            #%>%  filter(balance == "1"), 9å
              aes(x = length_sc, y = silver_vis, color = factor(balance))) +
  facet_wrap(~su)

```

#### L50 variation
```{r}

res_L50_90 <- lapply(1:nsu, function(j){
  
  dat_j <- df_data %>% filter(su == j)
  coef_j <- df_coef %>% filter(su == j)
    
  X_ind <- cbind(dat_j$age_sc, dat_j$con_index_sc, dat_j$eye_index_sc, dat_j$pec_index_sc, dat_j$habitat, dat_j$lat_sc)
  B_draws <- cbind(coef_j$bA, coef_j$bC, coef_j$bE, coef_j$bP, coef_j$bH, coef_j$bLa)
  
  # average covariates over individuals
  covar_eff <- rowMeans(B_draws %*% t(X_ind)) # average over individuals: n_draws vector
  
  # covariate and intercept contribution
  intercept_draw <- coef_j$alpha + covar_eff
  
  # L50: logit(0.5) = 0
  L50_sc <- (0 - intercept_draw) / exp(coef_j$logb0)
  
  # L90: logit(0.9) = log(0.9/0.1)
  L90_sc <- (log(0.9/0.1) - intercept_draw) / exp(coef_j$logb0)

  data.frame(
    su  = j,
    draw = 1:length(L50_sc),
    L50 = L50_sc,
    L90 = L90_sc
    )
  
  }) %>% bind_rows()

res_L50_90_long <- res_L50_90 %>%
  mutate(L50 = L50*sd(df.sweel3$length, na.rm = TRUE) + mean(df.sweel3$length, na.rm = TRUE),
         L90 = L90*sd(df.sweel3$length, na.rm = TRUE) + mean(df.sweel3$length, na.rm = TRUE)) %>% 
  pivot_longer(cols = c(L50,L90), names_to = "prob", values_to = "L") %>%
  left_join(df_suinfo)

res_L50_90_long %>%
  filter(su %in% c(1,4,13,54,60,61,23,43)) %>%
  ggplot() +
  geom_density(aes(L, fill = prob), color = NA, alpha= 0.3, show.legend = FALSE) +
  facet_wrap(~su, scales = "free_y")

```

### logistic p ~ age

#### age logistic curves
```{r}
# Adding covariate effects to the L50
# First make mean observed lat and inland/marine to each su
df.sweel3 %>%
  filter(age < 4) %>%
  ggplot() +
  geom_density(aes(length,color = factor(age)))
#ages <- seq(min(df.sweel3$age_sc, na.rm = TRUE), max(df.sweel3$age_sc, na.rm = TRUE), length.out = 100)
ages <- seq(-3, 7, length.out = 100)

res_A <- vector("list", nsu)

for(j in 1:nsu){
  
  dat_j <- df_data %>% filter(su == j)
  coef_j <- df_coef %>% filter(su == j)
  
  res_j <- lapply(ages, function(a){
    
    X_ind <- cbind(dat_j$length_sc,
                   rep(a, nrow(dat_j)),   # age_sc fixed to a
                   dat_j$con_index_sc,
                   dat_j$eye_index_sc, dat_j$pec_index_sc, dat_j$habitat, dat_j$lat_sc)
    
    B_draws <- cbind(exp(coef_j$logb0),coef_j$bA, coef_j$bC, coef_j$bE, coef_j$bP, coef_j$bH, coef_j$bLa)
    z_unit <- coef_j$alpha
    
    z_mat <- B_draws %*% t(X_ind) + as.vector(z_unit)  # n_draws x n_individuals
  
    p_mat      <- 1 / (1 + exp(-z_mat))
    p_marginal <- rowMeans(p_mat) # average over individuals: n_draws vector
  
    data.frame(
      su        = j,
      age_sc    = a,
      p_mean    = mean(p_marginal),
      p_lo      = quantile(p_marginal, 0.025),
      p_up      = quantile(p_marginal, 0.975)
      )
    })
  
  res_A[[j]] <- bind_rows(res_j)
}

# by habitat and sea basin
bind_rows(res_A) %>%
  #mutate(age = age_sc*sd(df.sweel3$age, na.rm = TRUE) + mean(df.sweel3$age, na.rm = TRUE)) %>% 
  left_join(df_suinfo) %>%
  ggplot() +
  geom_line(aes(x = age_sc, y = p_mean, group = lat_scmean, color = lat_scmean, alpha = balance)) +
  geom_point(data = df_inddra_mq %>% left_join(df.sweel3 %>% distinct(ind_id,silver_vis)) %>% left_join(df_suinfo %>% select(-sb)), 
             aes(age_sc, y = silver_vis, group = lat_scmean, color = lat_scmean, alpha = balance), show.legend = FALSE, alpha = 0.3, shape = 124, size = 1.5) +
  facet_wrap(~sea_basin) +
  scale_color_viridis_c()
ggsave(paste0(home, "/figures/L50_hab_basin.pdf"), width = 18, height = 13, units = "cm")

# by each spatial unit
bind_rows(res_A) %>%
  left_join(df_suinfo %>% select(-sb)) %>%
  ggplot() +
  geom_line(aes(age_sc, p_mean, color = sea_basin)) +
  geom_ribbon(aes(x = age_sc, y = p_mean, ymin = p_lo, ymax = p_up), alpha = 0.3, color = NA) +
  geom_point(data = df_inddra_mq %>% left_join(df.sweel3 %>% distinct(ind_id,length_sc,silver_vis)) %>% left_join(df_suinfo %>% select(-sb)),
  #            #%>%  filter(balance == "1"), 9å
              aes(x = length_sc, y = silver_vis, color = factor(balance))) +
  facet_wrap(~su)

```

#### A50 variation
```{r}

res_A50_90 <- lapply(1:nsu, function(j){
  
  dat_j <- df_data %>% filter(su == j)
  coef_j <- df_coef %>% filter(su == j)
    
  # X_ind <- cbind(dat_j$age_sc, dat_j$con_index_sc, dat_j$eye_index_sc, dat_j$pec_index_sc, dat_j$habitat, dat_j$lat_sc)
  # B_draws <- cbind(coef_j$bA, coef_j$bC, coef_j$bE, coef_j$bP, coef_j$bH, coef_j$bLa)
  X_ind <- cbind(dat_j$length_sc, dat_j$con_index_sc, dat_j$eye_index_sc, dat_j$pec_index_sc, dat_j$habitat, dat_j$lat_sc)
  B_draws <- cbind(coef_j$logb0, coef_j$bC, coef_j$bE, coef_j$bP, coef_j$bH, coef_j$bLa)
  
  # average covariates over individuals
  covar_eff <- rowMeans(B_draws %*% t(X_ind)) # average over individuals: n_draws vector
  
  # covariate and intercept contribution
  intercept_draw <- coef_j$alpha + covar_eff
  
  # L50: logit(0.5) = 0
  A50_sc <- (0 - intercept_draw) / exp(coef_j$bA)
  
  # L90: logit(0.9) = log(0.9/0.1)
  A90_sc <- (log(0.9/0.1) - intercept_draw) / exp(coef_j$bA)

  data.frame(
    su  = j,
    draw = 1:length(A50_sc),
    A50 = A50_sc,
    A90 = A90_sc
    )
  
  }) %>% bind_rows()

res_A50_90 %>%
  mutate(A50 = A50*sd(df.sweel3$length, na.rm = TRUE) + mean(df.sweel3$length, na.rm = TRUE),
         A90 = A90*sd(df.sweel3$length, na.rm = TRUE) + mean(df.sweel3$length, na.rm = TRUE)) %>% 
  pivot_longer(cols = c(A50,A90), names_to = "prob", values_to = "L") %>%
  filter(su %in% c(1,4,13,54,60,61,23,43)) %>%
  ggplot() +
  geom_density(aes(L, fill = prob), color = NA, alpha= 0.3, show.legend = FALSE) +
  facet_wrap(~su, scales = "free_y")
```

### logistic p ~ weight

#### condition logistic curves

```{r}
# Adding covariate effects to the L50
# First make mean observed lat and inland/marine to each su
cons <- seq(min(df.sweel3$con_index_sc, na.rm = TRUE), max(df.sweel3$con_index_sc, na.rm = TRUE), length.out = 100)
res_c <- vector("list", nsu)

for(j in 1:nsu){
  
  dat_j <- df_data %>% filter(su == j)
  coef_j <- df_coef %>% filter(su == j)
  
  res_j <- lapply(cons, function(c){
    
    X_ind <- cbind(dat_j$length_sc,dat_j$age_sc,
                   rep(c, nrow(dat_j)),   # con_index_sc fixed to c
                   dat_j$eye_index_sc, dat_j$pec_index_sc, dat_j$habitat, dat_j$lat_sc)
    
    B_draws <- cbind(exp(coef_j$logb0),coef_j$bA, coef_j$bC, coef_j$bE, coef_j$bP, coef_j$bH, coef_j$bLa)
    #z_unit <- coef_j$alpha + exp(coef_j$logb0)*l
    #z_unit <- coef_j$alpha + exp(coef_j$logb0) * dat_j$length_sc  # vector, one per individual
    z_unit <- coef_j$alpha
    
    z_mat <- B_draws %*% t(X_ind) + as.vector(z_unit)  # n_draws x n_individuals
  
    p_mat      <- 1 / (1 + exp(-z_mat))
    p_marginal <- rowMeans(p_mat) # average over individuals: n_draws vector
  
    data.frame(
      su        = j,
      con_index_sc = c,
      p_mean    = mean(p_marginal),
      p_lo      = quantile(p_marginal, 0.025),
      p_up      = quantile(p_marginal, 0.975)
      )
    })
  
  res_c[[j]] <- bind_rows(res_j)
}

# by habitat and sea basin
bind_rows(res_c) %>%
  left_join(df_suinfo) %>%
  #mutate(con_index = con_index_sc*sd(df.sweel3$con_index, na.rm = TRUE) + mean(df.sweel3$con_index, na.rm = TRUE)) %>%
  ggplot() +
  geom_line(aes(x = con_index_sc, y = p_mean, group = lat_scmean, color = lat_scmean, alpha = factor(balance))) +
  #geom_point(data = dfs %>% left_join(df.sweel3 %>% distinct(ind_id,length_sc,silver_vis)) %>% left_join(df_suinfo %>% select(-sb)), aes(length, y = silver, group = lat_scmean, color = lat_scmean, alpha = balance), show.legend = FALSE, alpha = 0.3, shape = 124, size = 1.5) +
  facet_wrap(~sea_basin) +
  scale_color_viridis_c()
ggsave(paste0(home, "/figures/w50_hab_basin.pdf"), width = 18, height = 13, units = "cm")

# by each spatial unit
# bind_rows(res_c) %>%
#   #mutate(length = length_sc*sd(df.sweel3$length, na.rm = TRUE) + mean(df.sweel3$length, na.rm = TRUE)) %>% 
#   left_join(df_suinfo %>% select(-sb)) %>%
#   #filter(balance == "1") %>%
#   ggplot() +
#   geom_line(aes(length_sc, p_mean, color = sea_basin)) +
#   geom_ribbon(aes(x = length_sc, y = p_mean, ymin = p_lo, ymax = p_up), alpha = 0.3, color = NA) +
#   #, group = spat_unit), fill = factor(balance)), alpha = 0.5, color = NA) + 
# #geom_point(data = df.sweel3, aes(length, y = silver_vis)) +
#   geom_point(data = dfs %>% left_join(df.sweel3 %>% distinct(ind_id,length_sc,silver_vis)) %>% left_join(df_suinfo %>% select(-sb)),
#   #            #%>%  filter(balance == "1"), 9å
#               aes(x = length_sc, y = silver_vis, color = factor(balance))) +
#   facet_wrap(~su)

```

#### C50 variation
```{r}
res_C50_90 <- lapply(1:nsu, function(j){
  
  dat_j <- df_data %>% filter(su == j)
  coef_j <- df_coef %>% filter(su == j)
    
  # X_ind <- cbind(dat_j$age_sc, dat_j$con_index_sc, dat_j$eye_index_sc, dat_j$pec_index_sc, dat_j$habitat, dat_j$lat_sc)
  # B_draws <- cbind(coef_j$bA, coef_j$bC, coef_j$bE, coef_j$bP, coef_j$bH, coef_j$bLa)
  X_ind <- cbind(dat_j$length_sc, dat_j$age_sc, dat_j$eye_index_sc, dat_j$pec_index_sc, dat_j$habitat, dat_j$lat_sc)
  B_draws <- cbind(coef_j$logb0, coef_j$bA, coef_j$bE, coef_j$bP, coef_j$bH, coef_j$bLa)
  
  # average covariates over individuals
  covar_eff <- rowMeans(B_draws %*% t(X_ind)) # average over individuals: n_draws vector
  
  # covariate and intercept contribution
  intercept_draw <- coef_j$alpha + covar_eff
  
  # L50: logit(0.5) = 0
  C50_sc <- (0 - intercept_draw) / exp(coef_j$bC)
  
  # L90: logit(0.9) = log(0.9/0.1)
  C90_sc <- (log(0.9/0.1) - intercept_draw) / exp(coef_j$bC)

  data.frame(
    su  = j,
    draw = 1:length(C50_sc),
    C50 = C50_sc,
    C90 = C90_sc
    )
  
  }) %>% bind_rows()

res_C50_90 %>%
  mutate(C50 = C50*sd(df.sweel3$con_index, na.rm = TRUE) + mean(df.sweel3$con_index, na.rm = TRUE),
         C90 = C90*sd(df.sweel3$con_index, na.rm = TRUE) + mean(df.sweel3$con_index, na.rm = TRUE)) %>% 
  pivot_longer(cols = c(C50,C90), names_to = "prob", values_to = "L") %>%
  filter(su %in% c(1,4,13,54,60,61,23,43)) %>%
  ggplot() +
  geom_density(aes(L, fill = prob), color = NA, alpha= 0.3, show.legend = FALSE) +
  facet_wrap(~su, scales = "free_y")

```

### Combine propability curves
```{r}
bind_rows(res_l) %>%
  rename(val = length_sc) %>%
  mutate(trait = "length") %>%
  bind_rows(bind_rows(res_A) %>%
              rename(val = age_sc) %>%
              mutate(trait = "age")) %>%
  bind_rows(bind_rows(res_c) %>%
              rename(val = con_index_sc) %>%
              mutate(trait = "con")) %>%
  left_join(df_suinfo) %>%
  filter(habitat == 0) %>%
  ggplot() +
  geom_line(aes(x = val, y = p_mean, group = lat_scmean, 
                color = lat_scmean, alpha = balance)) + 
  facet_wrap(~trait, scales = "free") +
  guides(alpha = FALSE)  
    
res_L50_90 %>%
  mutate(L50 = L50*sd(df.sweel3$length, na.rm = TRUE) + mean(df.sweel3$length, na.rm = TRUE),
         L90 = L90*sd(df.sweel3$length, na.rm = TRUE) + mean(df.sweel3$length, na.rm = TRUE)) %>% 
  pivot_longer(cols = c(L50,L90), names_to = "prob", values_to = "val") %>%
  mutate(trait = "length") %>%
  bind_rows(res_A50_90 %>%
              mutate(A50 = A50*sd(df.sweel3$age, na.rm = TRUE) + mean(df.sweel3$age, na.rm = TRUE),
                     A90 = A90*sd(df.sweel3$age, na.rm = TRUE) + mean(df.sweel3$age, na.rm = TRUE)) %>%
              pivot_longer(cols = c(A50,A90), names_to = "prob", values_to = "val") %>%
              mutate(trait = "age")) %>%
  bind_rows(res_C50_90 %>%
              mutate(C50 = C50*sd(df.sweel3$con_index, na.rm = TRUE) + mean(df.sweel3$con_index, na.rm = TRUE),
                     C90 = C90*sd(df.sweel3$con_index, na.rm = TRUE) + mean(df.sweel3$con_index, na.rm = TRUE)) %>% 
              pivot_longer(cols = c(C50,C90), names_to = "prob", values_to = "val") %>% 
              mutate(trait = "condition")) %>%
  #filter(su %in% c(1,4,13,54,60,61,23,43)) %>%
  ggplot() +
  geom_density(aes(val, fill = prob, group =su), color = NA, alpha= 0.3, show.legend = FALSE) +
  facet_wrap(~prob, ncol = 2, scales = "free")
```


### Data: Length, Age and Weight

#### In space
```{r, cache = FALSE}
# Catchment polygons
ca_sf <-readRDS(paste0(home,"/R/prep_data/catchment_polygons.rds"))
ba_sf <-readRDS(paste0(home,"/R/prep_data/basin_polygons.rds"))
ha_sf <-readRDS(paste0(home,"/R/prep_data/haro_polygons.rds"))

lht_har <- df_inddra_mq %>%
  filter(silver == 1) %>%
  left_join(df.sweel3 %>%
              select(length,age,weight,ind_id)) %>%
  summarise(ml = median(length),
            al = median(age,na.rm= TRUE), 
            wl = median(weight,na.rm= TRUE), .by = spat_unit) %>%
  left_join(ha_sf %>% rename(spat_unit = haro)) %>%
  filter(!is.na(ogc_fid))

lht_cat <- df_inddra_mq %>%
  filter(silver == 1) %>%
  left_join(df.sweel3 %>%
              select(length,age,weight,ind_id)) %>%
  summarise(ml = median(length),
            al = median(age,na.rm= TRUE), 
            wl = median(weight,na.rm= TRUE), .by = spat_unit) %>%
  left_join(ca_sf %>% mutate(spat_unit = as.character(main_catch))) %>%
  filter(!is.na(main_catch))
  
lht_bas <- df_inddra_mq %>%
  filter(silver == 1) %>%
  left_join(df.sweel3 %>%
              select(length,age,weight,ind_id)) %>%
  summarise(ml = median(length),
            al = median(age,na.rm= TRUE), 
            wl = median(weight,na.rm= TRUE), .by = spat_unit) %>%
  left_join(ba_sf %>% mutate(spat_unit = sea_basin)) %>%
  filter(!is.na(sea_basin))


pl <- plot_map_Nord +
  geom_sf(data = lht_bas %>% st_as_sf(), aes(fill = ml), color = NA) +
  geom_sf(data = lht_cat %>% st_as_sf(), aes(fill = ml), color = NA) +
  geom_sf(data = lht_har %>% st_as_sf(), aes(fill = ml), color = NA) +
  scale_fill_viridis_c(option = "C",limits=c(min(space_lht$ml,na.rm = TRUE),max(space_lht$ml,na.rm = TRUE))) +
  theme_minimal() +
  labs(fill = "Length [mm]") +
  theme(
    axis.title.x = element_blank(),
    legend.position = c(0.17, 0.75),
    legend.key.width = unit(0.3, "cm"),
    legend.key.height = unit(0.3, "cm"),
    legend.title = element_text(size = 8))

pa <- plot_map_Nord +
  geom_sf(data = lht_bas %>% st_as_sf() %>% drop_na(al), aes(fill = al), color = NA) +
  geom_sf(data = lht_cat %>% st_as_sf() %>% drop_na(al), aes(fill = al), color = NA) +
  geom_sf(data = lht_har %>% st_as_sf() %>% drop_na(al), aes(fill = al), color = NA) +
  scale_fill_viridis_c(option = "C",limits=c(min(space_lht$al, na.rm = TRUE),max(space_lht$al,na.rm = TRUE))) +
  theme_minimal() +
  labs(fill = "Age [years]") +
  theme(
    legend.position = c(0.17, 0.75),
    legend.key.width = unit(0.3, "cm"),
    legend.key.height = unit(0.3, "cm"),
    legend.title = element_text(size = 8)
  )

pw <- plot_map_Nord +
  geom_sf(data = lht_bas %>% st_as_sf() %>% drop_na(wl), aes(fill = wl), color = NA,  alpha = 0.7) +
  geom_sf(data = lht_cat %>% st_as_sf() %>% drop_na(wl), aes(fill = wl), color = NA,  alpha = 0.8) +
  geom_sf(data = lht_har %>% st_as_sf() %>% drop_na(wl), aes(fill = wl), color = NA,  alpha = 0.8) +
  scale_fill_viridis_c(option = "C",limits=c(min(space_lht$wl, na.rm = TRUE),max(space_lht$wl, na.rm = TRUE))) +
  theme_minimal() +
  labs(fill = "Weight [g]") +
  theme(
    legend.position = c(0.17, 0.75),
    legend.key.width = unit(0.3, "cm"),
    legend.key.height = unit(0.3, "cm"),
    legend.title = element_text(size = 8)
  ) +
  theme(axis.title.x=element_blank())
 
pl + pw + pa
```

#### by su and over space
```{r, cache = FALSE}
# probability vs silver
# df_inddra_mq %>%
#   left_join(df.sweel3 %>% 
#                select(length,ind_id)) %>%
#   ggplot() +
#   geom_jitter(aes(y = silver, x = p, color = length), alpha = 0.7, position = "jitter", size = 0.3) 


# Weighting by p has very little effect on the density distribution:
pl2 <- df_inddra_mq %>%
  filter(silver == 1) %>%
  left_join(df.sweel3 %>% 
               select(length,ind_id)) %>%
  mutate(ml = median(length), .by = spat_unit) %>%
  ggplot(aes(x = length, y = reorder(spat_unit, ml))) + 
  geom_density_ridges(aes(fill = ml),
                      color = NA, show.legend = FALSE, alpha = 0.8) +
  #geom_density_ridges(aes(weight = p, color = ml),
  #                    fill = NA) ?
  scale_fill_viridis_c(name = "Median length [mm]", option = "C", limits=c(min(df_ind_su$ml),max(df_ind_su$ml))) +
  geom_text(data = . %>% summarise(n = n(), ml = median(length), .by = spat_unit),
            aes(x = 1250, label = paste0("n = ", n)),
            hjust = 1, vjust = -0.2, size = 3, color = "darkgrey") +
  labs(y = "Spatial unit", x = "Length [mm]") +
  xlim(400,1250) +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.y=element_blank())

pa2 <- df_inddra_mq %>%
  filter(silver == 1) %>%
  left_join(df.sweel3 %>% 
               select(age,length,ind_id)) %>%
  drop_na(age) %>%
  mutate(ml = median(length),
         al = median(age,na.rm = TRUE), .by = spat_unit) %>%
  ggplot(aes(x = age, y = reorder(spat_unit, ml))) + 
  geom_density_ridges(aes(fill = al),
                      color = NA, show.legend = FALSE, alpha = 0.8) +
  scale_fill_viridis_c(name = "Median age [years]", option = "C", limits=c(min(df_ind_su$al),max(df_ind_su$al))) +
  geom_text(data = . %>% summarise(n = n(), ml = median(length), .by = spat_unit),
            aes(x = 35, label = paste0("n = ", n)),
            hjust = 1, vjust = -0.2, size = 3, color = "darkgrey") +
  labs(y = "Spatial unit", x = "Age [years]") +
  xlim(NA,35) +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.y=element_blank()) 

pw2 <- df_inddra_mq %>%
  filter(silver == 1) %>%
  left_join(df.sweel3 %>% 
               select(weight,length,ind_id)) %>%
  drop_na(weight) %>%
  mutate(ml = median(length),
         wl = median(weight,na.rm = TRUE), .by = spat_unit) %>%
  ggplot(aes(x = weight, y = reorder(spat_unit, ml))) + 
  geom_density_ridges(aes(fill = wl),
                      color = NA, show.legend = FALSE, alpha = 0.8) +
  #geom_density_ridges(aes(weight = p, color = ml),
  #                    fill = NA) ?
  scale_fill_viridis_c(name = "Median weight [g]", option = "C", limits=c(min(df_ind_su$wl),max(df_ind_su$wl))) +
  geom_text(data = . %>% summarise(n = n(), ml = median(length), .by = spat_unit),
            aes(x = Inf, label = paste0("n = ", n)),
            hjust = 1, vjust = -0.2, size = 3, color = "darkgrey") +
  labs(y = "Spatial unit", x = "Age [years]") +
  labs(y = "Spatial unit", x = "Weight [g]") +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.y=element_blank())
```


```{r, cache = FALSE}
df_inddra_mq %>%
  filter(marine == "no",
         silver == 1) %>%
  left_join(df.sweel3 %>% 
               select(length,ind_id)) %>%
  mutate(ml = median(length), .by = spat_unit) %>%
  ggplot(aes(x = length, y = factor(spat_unit))) + 
  geom_density_ridges(aes(fill = ml),
                    color = NA, show.legend = FALSE, alpha = 0.8) +
  geom_density_ridges(data = res_L50_90_long %>% 
                        mutate(ml = median(L), .by = spat_unit) %>%
                        filter(habitat == 0,
                               spat_unit %in% c(df_inddra_mq %>% 
                                 filter(marine == "no", silver == 1) %>% distinct(spat_unit) %>% pull())),
                      aes(y = factor(spat_unit), x = L, color = prob), fill = NA) +
  scale_fill_viridis_c(name = "Median length [mm]", option = "C", limits=c(min(df_ind_ca$ml),max(df_ind_ca$ml))) +
  #scale_color_viridis_c(name = "Median length [mm]", option = "C", limits=c(min(df_ind_ca$ml),max(df_ind_ca$ml))) +
  labs(y = "Spatial unit", x = "Length [mm]") +
  xlim(400,1270) +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) 
# df_inddra_mq %>%
#   filter(marine == "no",
#          silver == 1) %>%
#   left_join(df.sweel3 %>% 
#                select(length,ind_id)) %>%
#   mutate(ml = median(length), .by = spat_unit) %>%
#   ggplot(aes(x = length, y = factor(sb))) + 
#   geom_density_ridges(aes(fill = ml),
#                       bandwidth = 1.5,
#                       color = NA, show.legend = FALSE, alpha = 0.3) +
#   #y = reorder(spat_unit, ml)
#   geom_density_ridges(aes(weight = p, color = ml),
#                       bandwidth = 1.5,
#                       fill = NA) 
#     scale_fill_viridis_c(name = "Median length [mm]", option = "C", limits=c(min(df_ind_ca$ml),max(df_ind_ca$ml))) +
#   scale_color_viridis_c(name = "Median length [mm]", option = "C", limits=c(min(df_ind_ca$ml),max(df_ind_ca$ml))) +
#   labs(y = "Spatial unit", x = "Length [mm]") +
#   xlim(400,1270) +
#   theme(axis.text.y=element_blank(),
#         axis.ticks.y=element_blank()) 

df_inddra_mq %>%
  filter(marine == "no",
         silver == 1) %>%
  left_join(df.sweel3 %>% 
               select(length,ind_id)) %>%
  mutate(ml = median(length), .by = spat_unit) %>%
  ggplot(aes(x = length)) + 
  stat_density(aes(weight = p, fill = ml), 
               position = "identity", 
               alpha = 0.5, 
               geom = "area") +
  stat_density(aes(color = ml), 
               position = "identity", 
               fill = NA, 
               linewidth = 0.5, 
               geom = "line") +
  facet_grid(spat_unit ~ ., scales = "free_y") +
  #facet_grid(reorder(spat_unit, ml) ~ ., scales = "free_y") +
  scale_fill_viridis_c(name = "Median length [mm]", option = "C", limits=c(min(df_ind_ca$ml),max(df_ind_ca$ml))) +
  scale_color_viridis_c(name = "Median length [mm]", option = "C", limits=c(min(df_ind_ca$ml),max(df_ind_ca$ml))) +
  theme_minimal() +
  labs(y = "Spatial unit", x = "Length [mm]") +
  xlim(400,1200) +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) 

res_L50_90_long %>%
  mutate(ml = median(L), .by = spat_unit) %>%
  filter(habitat == 0,
         spat_unit %in% c(df_inddra_mq %>% 
                            filter(silver == 1) %>% 
                            distinct(spat_unit) %>% pull())) %>%
  ggplot(aes(x = L, y = factor(spat_unit))) + 
  geom_density_ridges(aes(color = ml, fill = prob),
                    color = NA, show.legend = FALSE, alpha = 0.8) +
  geom_boxplot(data = df_inddra_mq %>% filter(silver == 1) %>%
               left_join(df.sweel3 %>% select(length,ind_id)),
               aes(x = length, y = factor(spat_unit)), alpha = 0.1)

```

### Combine by-SU plots
```{r}
# indiviudally
pl3 <- pl+ pl2 + plot_layout(guides = "collect")
ggsave(paste0(home, "/figures/meanlength.pdf"), width = 17, height = 14, units = "cm")
pa3 <- pa + pa2 + plot_layout(guides = "collect")
ggsave(paste0(home, "/figures/meanage.pdf"), width = 17, height = 14, units = "cm")
pw3 <- pw + pw2 + plot_layout(guides = "collect")
ggsave(paste0(home, "/figures/meanweight.pdf"), width = 17, height = 14, units = "cm")
```


```{r}
#| fig-width: 7 
#| fig-height: 12
# combine all
pl + pl2 + pw + pw2 + pa + pa2 + plot_layout(nrow = 3)
ggsave(paste0(home, "/figures/meanall.pdf"), width = 17, height = 24, units = "cm")

```

### Posterior predictive checks
```{r}
# visual classifier
# compare means by spatial unit
tmps <- as_draws(s.samples) %>%
  #subset_draws(variable = c("visual_rep")) %>%
  spread_draws(visual_pp[ind_id],silver_pp[ind_id],silver[ind_id]) %>%
  median_qi() %>%
  left_join(df.sweel3 %>% select(su,spat_unit,ind_id,silver_vis,durif_stage), by = "ind_id") 

tmps %>%
  summarise(mean_est_vis = mean(visual_pp, na.rm = TRUE), .by = spat_unit) %>%
  left_join(df.sweel3 %>% summarise(obs_vis = mean(silver_vis, na.rm = TRUE), .by = spat_unit), by = "spat_unit") %>%
#  left_join(df_suinfo) %>%
  ggplot() +
  geom_point(aes(obs_vis,mean_est_vis)) +
  geom_abline() 

tmps %>%
  left_join(df.sweel3 %>% select(durif_stage,ind_id,silver_vis,length_sc)) %>%
  summarise(n = n(), ml = mean(length_sc), .by = c(durif_stage,visual_pp)) %>%
  arrange(-n) 

tmps %>%
  left_join(df.sweel3 %>% select(durif_stage,ind_id,silver_vis,length)) %>%
  summarise(n = n(), ml = mean(length), .by = c(silver_vis,silver_rep,visual_rep)) %>%
  arrange(-n) 

left_join(df_suinfo) %>%
  ggplot() +
  geom_point(aes(durif_stage,visual_rep)) 

# plot densities of 0/1 
tmps %>%
  ggplot() +
  geom_density(aes(visual_rep), color = "red") +
  geom_density(data = df.sweel3, aes(silver_vis)) +
  labs(title = "red is model") 

tmps %>%
  summarise(mean_est_vis = sum(visual_pp, na.rm = TRUE), .by = spat_unit) %>%
  left_join(df.sweel3 %>% summarise(obs_vis = sum(silver_vis, na.rm = TRUE), .by = spat_unit), by = "spat_unit") %>%
#  left_join(df_suinfo) %>%
  ggplot() +
  geom_point(aes(obs_vis,mean_est_vis, color = factor(balance))) +
  geom_abline()

# for the balancing by inroducing NAs model
tmps2 <- df.sweel3 %>%
  mutate(pre_silver_vis = case_when(visual_stage == "Blankål" ~ 1,
                                visual_stage == "Halvblank ål" ~ NA,
                                visual_stage %in% c("Gulål","Yngel","Juvenil") ~ 0,
                                .default = NA)) %>%
  select(silver_vis,pre_silver_vis,ind_id) %>%
  left_join(as_draws(s.samples) %>%
              spread_draws(silver[ind_id]) %>%
              median_qi()) 
tmps2 %>%
  # select those that were set to NA before model
  filter(is.na(pre_silver_vis) == !is.na(silver_vis)) %>%
  summarise(n(), .by = c(silver,pre_silver_vis)) 
9238/(5739+9283)
2733/(4812+2733)


# Durif classifier
# tmpd <- as_draws(s.samples) %>%
#   #subset_draws(variable = c("visual_rep")) %>%
#   spread_draws(measured_rep[ind.id]) %>%
#   median_qi() %>%
#   left_join(df.sweel3 %>% select(su,spat_unit,ind.id), by = "ind.id") 
# 
# tmpd %>%
#   summarise(mean_est_dur = mean(measured_rep, na.rm = TRUE), .by = spat_unit) %>%
#   left_join(df.sweel3 %>% summarise(obs_dur = mean(silver_dur, na.rm = TRUE), .by = spat_unit), by = "spat_unit") 
# 
# tmpd %>%
#   ggplot() +
#   geom_density(aes(measured_rep), color = "red") +
#   geom_density(data = df.sweel3, aes(silver_vis)) +
#   labs(title = "red is model")
#   ggplot() +
#   geom_density(aes(measured_rep)) +
#   geom_density(data = df.sweel3, aes(silver_dur))

# Test
# Disagreement rate between classifiers — the model's ability to reproduce the observed rate of V≠M is a direct check on Se/Sp estimation
#Prevalence by population — check that simulated stage prevalence matches observed classifier prevalence within each population mb
#Prevalence along the length gradient — bin individuals by length and compare observed vs simulated classifier rates, which checks the logistic curve f

```

### Predicted / observed

```{r}
# var.lats <- data.silv %>%
#   mutate(m.lat = mean(ser_y), .by = ecoreg) %>%
#   distinct(er, ecoreg, m.lat) %>%
#   arrange(-m.lat)

dfp <- s.samples %>%
  as_draws() %>%
  spread_draws(mu_s[ind.id]) %>%
  median_qi() %>%
  rename(pred.sl = mu_s) %>%
  ungroup() %>%
  left_join(df.sweel %>%
              select(main_bas,length,lat,lon,sex,ind.id) %>%
              rename(obs.sl = length))
  #left_join(var.lats) %>%
  #mutate(ecoreg = fct_reorder(as.factor(ecoreg), -m.lat))

dfp %>%
  #filter(is.na(sex)) %>%
  ggplot() +
  geom_abline(slope = 1, color = "darkgrey") +
  geom_point(aes(x = obs.sl, y = exp(pred.sl), color = sex), shape = 4, alpha = 0.7, show.legend = FALSE) +
  facet_wrap(~main_bas) +
  theme_minimal() +
  labs(y="predicted", x="observed")
#ggsave(paste0(home, "/report/silv_po.png"), width = 17, height = 14, units = "cm")

```


### The durifs

```{r}
tmp3s <- s.samples %>%
  as_draws() %>%
  spread_draws(silver[ind_id]) %>%
  median_qi() %>%
  left_join(df.sweel3 %>% select(visual_stage,durif_stage,ind_id,su,balance,length), by = "ind_id")

tmp3s %>%
  filter(durif_stage %in% c("FII","FIII","FIV","FV")) %>%
  summarise(n = n(), .by = c(silver, durif_stage, visual_stage)) %>%
  arrange(-silver, -n) %>%
  rename(estimated = silver)


```

### Compare visual and durif with classifier

```{r}
tmp4 <- s.samples %>%
  as_draws() %>%
  spread_draws(silver[ind_id],p[ind_id],p_vis[ind_id]) 

tmp4b <- tmp4 %>%
  group_by(ind_id) %>%
  median_qi() %>%
  left_join(df.sweel3 %>% select(silver_vis,visual_stage,durif_stage,ind_id,su,length,balance), by = "ind_id")

df_inddra_mq %>%
  ggplot(aes(x=length,y=p)) +
  geom_point( alpha =0.1) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 3), show.legend = FALSE) +
  facet_wrap(~silver_vis)
# tmp4b <- tmp4 %>%
#   filter(silver == 1,
#          balance == 1) %>%
#   summarise(mean = mean(length),
#             sd = sd(length), .by = su) %>%
#   ungroup() %>%
#   cbind(tmp4 %>%
#           filter(visual_stage == "Blankål",
#                  balance == 1) %>%
#           summarise(meanv = mean(length),
#                     sdv = sd(length), .by = su) %>%
#           ungroup() %>%
#           select(-su))

tmp4b %>%
  filter(p > 0.9,
         silver == 0)
1-146/66207 # sensitivty sort of!

df_inddra_mq_c <- df_inddra_mq %>%
  left_join(df.sweel3 %>% select(length, ind_id)) %>%
  filter(p_vis > 0.9,
         balance == 1) %>%
  #summarise(mean = mean(length),
  #          sd = sd(length), .by = su) %>%
  ungroup() %>%
  mutate(source = "model_pvis") %>%
  bind_rows(df_inddra_mq %>%
              left_join(df.sweel3 %>% select(length, ind_id)) %>%
              filter(p > 0.9,
                     balance == 1) %>%
              #summarise(mean = mean(length),
              #          sd = sd(length), .by = su) %>%
              ungroup() %>%
              mutate(source = "model_p")) %>%
  bind_rows(df_inddra_mq %>%
              left_join(df.sweel3 %>% select(length, ind_id,visual_stage)) %>%
              filter(visual_stage == "Blankål",
                     balance == 1) %>%
              #summarise(mean = mean(length),
              #          sd = sd(length), .by = su) %>%
              ungroup() %>%
              mutate(source = "visual")) %>%
  bind_rows(df_inddra_mq %>%
              left_join(df.sweel3 %>% select(length, ind_id,durif_stage)) %>%
              filter(durif_stage %in% c("FIII","FIV","FV"),
                     balance == 1) %>%
              #summarise(mean = mean(length),
              #          sd = sd(length), .by = su) %>%
              ungroup() %>%
              mutate(source = "Durif"))


# df_inddra_mq_c %>%
#   expand_grid(length = seq(400, 1400, length.out = 300)) %>%
#   mutate(y = dnorm(length, mean = mean, sd = sd)) %>%
# #  filter(su %in% c(15)) %>%
#   ggplot() +
#   #geom_area(aes(length, y, fill = source), position = "identity", alpha = 0.4) +
#   geom_line(aes(length, y,  color = source)) +
#   facet_wrap(~su) +
#   geom_vline(xintercept = mean(df.sweel3$length), color = "darkgrey") 
df_inddra_mq_c %>%
  ggplot() +
  geom_boxplot(aes(length, y = source,  color = source),
               outlier.size = 0.5,
               outlier.alpha = 0.5,
               staple.linewidth = 0.5) +
  facet_wrap(~su) +
  geom_vline(xintercept = mean(df.sweel3$length), color = "darkgrey")
ggsave(paste0(home, "/figures/compare_classif.pdf"), width = 18, height = 13, units = "cm")

# tmp5 %>%
#   filter(silver == 1) %>%
#   ggplot() +
#   geom_density(aes(length)) +
#   geom_vline(data = tmp5 %>% filter(silver == 1) %>% mutate(m = mean(length), .by = su), aes(xintercept = m)) +
#   geom_density(data = tmp5 %>% filter(visual_stage == "Blankål"), aes(length), color = "red") +
#   geom_vline(data = tmp5 %>% filter(visual_stage == "Blankål") %>% mutate(m = mean(length), .by = su), aes(xintercept = m), color = "red") +
#   geom_density(data = tmp5 %>% filter(durif_stage %in% c("FIII","FIV","FV")), aes(length), color = "blue") +
#   geom_vline(data = tmp5 %>% filter(durif_stage %in% c("FIII","FIV","FV")) %>% mutate(m = mean(length), .by = su), aes(xintercept = m), color = "blue") +
#   facet_wrap(~su)

tmp4b %>%
  mutate(p_diff = p_vis - p) %>%
  filter(abs(p_diff) > 0.5) %>%
  #arrange(-p_diff)
  summarise(n(), .by = silver)
  
```

# test distributions
```{r}
mu_a_gl ~ dnorm(msa, sda)
sd_a_gl ~ dgamma(shape = 1.5, rate = 1.5 / sda)
#mu_b_gl ~ dnorm(log(mb0), sd = log(sdb0))
mu_b_gl ~ dnorm(log(mb0), sd = 0.5)
#sd_b_gl ~ dgamma(shape = 1.5, rate = 1.5 / sdb0) # to weak 
sd_b_gl ~ dgamma(shape = 2, rate = 4 / sdb0) 

sd_a_su ~ dgamma(shape = 2, rate = 2cx)   # sd for su-level
  

ggplot() +
  #stat_function(fun = dgamma, 
  #            args = list(shape = 1000, rate = 2000)) +
  stat_function(fun = dgamma, 
              args = list(shape = 2,rate=4), color = "blue") + xlim(0,5)  
ggplot() +
  #stat_function(fun = dgamma, 
  #            args = list(shape = 1000, rate = 2000)) +
  stat_function(fun = dbeta, 
              args = list(5,2), color = "blue") 
  stat_function(fun = dexp, 
              args = list(7), color = "blue") +
  stat_function(fun = dexp, 
              args = list(3)) +
  xlim(0,5)
```

```{r}
tmpab <- 
  subset_draws(as_draws(s.samples), variable = c("ab")) %>%
  summarise_draws(c("mean","sd")) %>%
  mutate(variable = case_when(variable == "ab[1]" ~ "a",
                              variable == "ab[2]" ~ "b")) %>%
  pivot_wider(names_from = variable, values_from = c(mean,sd))
                           
df.sweel3 %>%
  select(length,weight,su) %>%
  drop_na(length,weight) %>%
  bind_cols(tmpab) %>%
  mutate(logweight_mu = mean_a + mean_b*log(length),
         le_cren = weight / exp(logweight_mu)) %>%
  filter(le_cren < 0.001) %>%
  ggplot() +
  #geom_point(aes(length, weight))
  geom_density(aes(le_cren
                  ))
```

